### 2009-03-30 text/paper.tex
Added paper stub.

\documentclass[12pt]{article}
\usepackage[utf8x]{inputenc}
\author{Magdalena Leshtanska \and Aleksandar Dimitrov}
\title{Error-Centric Annotation of Learner Corpora}
\begin{document}
\maketitle
\section{Introduction}\label{sec:intro}

Literature overview. Bla Bla.

All of the annotation schemes mentioned above focus on the \textbf{text} itself,
and will often even advise the annotator to \textit{modify} the underlying text
with the annotation. We believe that a more error-centric annotation of the data
can be beneficial.

Firstly, every annotation scheme that operates directly on the text or spans
over a piece of text will run into two kinds of problems:

\begin{itemize}
\item \textit{Interleaving annotations} occur when an error doesn't end before
another one begins. Given the tokens $a_1 a_2 a_3$, and two errors $e_1$ and
$e_2$ ranging over tokens $a_1$, $a_2$ and $a_2$ $a_3$ respectively, the
resulting markup will be confusing or outright impossible to read: $(e_1)a_1
(e_2)a_2(/e_1)a_3(/e_2)$. This is particularly a problem with XML,
since the specification\footnotemark explicitly disallows interleaving
markup.
\item \textit{Greedy annotation} covers tokens entirely uninvolved in the
\textit{``production''} of an error. If of the tokens $a_1 a_2 a_3$ only $a_1$
and $a_3$ are involved in the production of an error (say, an agreement error)
marking the entire token sequence as erroneous would involve blaming the
otherwise completely innocent token $a_2$. This can significantly increase the
noise in a corpus.
\end{itemize}
\footnotetext{Located at \texttt{http://www.w3.org/TR/REC-xml/}}

There is little data about the primary use cases for learner corpora known to
us, but it does not seem too far fetched to assume most linguists interested in
such data are not likely to browse through the corpora in order to find an
amusingly written piece of text. Instead, it seems to be safe to assume that
most would be interested in the \textbf{kinds of errors} that can occur in such
a corpus, and their particular properties. The error data should thus contain as
little noise as possible and be maximally specific while retaining a large amount
of generality.  It should furthermore be easily accessible, without having to
actually read the texts within the corpus itself.

Based on these assumptions, we decided to decouple the \textit{error markup}
from the \textit{corpus data}. Specifically, to our annotation scheme, the
errors and the text are two entirely different data structures. Every
particular error can reference tokens within the corpus using a \textbf{key},
similar to the way modern Relational Data Bases reference their data. This makes
the index more accessible and easier to maintain. Every marked up error contains exactly
the amount of information it needs in order to justify its presence, viz. by
defining what triggered its occurrence, and sufficient classification within the
taxonomy presented in \ref{sec:taxo}.

\section{Error Taxonomy and Annotation Scheme}\label{sec:scheme}

Learner errors are not easily classified due to their wide variety of
appearance, kind, and ambiguities involved. Typically, an annotation scheme will
strive to cover as large a scope it possibly can unambiguously classify, yet
there are many practical limitations to the potential coverage one can achieve.

One of the most prevalent restrictions is the need to devise the annotation
manual with great attention to detail, and to perform a meaningful assessment of
the quality of the annotated data after the annotation process. Even a most
meticulously well-defined manual is still going to be read and applied by
humans, who all have their personal intuition and character. It is easy to see
why this inherent individuality poses problems to the consistency of any kind of
annotated data. A well-written annotation manual can limit the effect this
natural variance can have on the data, but applying measures of consistency is
strictly necessary. The assessment of our data will be discussed in section
\ref{sec:results} while the remainder of this section will explain the rationale
behind our annotation manual, which is in turn given in section \ref{sec:manual}.

\subsection{Calling Errors by Name}\label{sec:taxo}

In order to identify errors occurring in learner language, one has to first
classify them. However, even then, identifying which class a particular error
belongs to is in no way an easy job. Devising an exhaustive taxonomy of errors
that can appear in natural language seems a daunting task, since science has so
far failed to regularize what \textit{is} a valid utterance of a language.
Therefore, our taxonomy will try to do some things only, and do them well. In
particular, we will not cover punctuation mistakes, since their possible
corrections can have cascading effects on the rest of the sentences and suddenly
give rise to all kinds of new errors.

Furthermore, our taxonomy will have a base case, called \textbf{gibberish} for
every error that cannot otherwise be classified, necessitated by its
deliberately established incompleteness. Sections delineated as such a case
should, however, be few in number, and as local as possible.

There are three other kinds of errors we want to term:

\begin{enumerate}
\item \textit{Spelling errors} are a primitive form of error. They contain no
meta-data or any kind of further specification except maybe a target word they
should have ended up as.
\item \textit{Form errors} are otherwise grammatical sentences which are,
however, semantically incorrect given the current context.
\item \textit{Grammar errors} simply do not belong to the target language because
of their erroneous morphology or syntax.
\end{enumerate}

\subsection{Representing Errors in Annotations}

\section{Annotation Manual}\label{sec:manual}

\subsection{Spelling Errors}\label{sec:speling}
A spelling error is an error that \textit{cannot} be explained by mistakes in
morphology, e.g. derivation or inflection. It will typically comprise only one
word. If it comprises more than one word, it is seen as a \textit{compound}
spelling error, i.e. two tokens separated by a space should have been
concatenated to yield only one token. Similarly, if the target hypothesis
accounts for more than one word, the token should have been split up into two.

\subsection{Form Errors}
Form errors occur in otherwise completely grammatical sentences, which, however
do not convey what was obviously intended by the author taking context into
account. Lexical errors are also form errors, if the given word is grammatically
valid within the sentence.

\subsection{Grammatical Errors}

\subsection{Devising Target Hypotheses}
Retain existing input by the author of the text to the greatest extent possible.
The text should be only minimally altered in the target hypothesis.

\section{Results, Assessment and Future Work}\label{sec:results}

\end{document}

### 2009-03-30 src/Annotator/DTD.hs
Intorduced some relevant taxonomy.

@@ -24,21 +24,24 @@ data Type = TypeForm Form
           | TypeGrammar Grammar
           | TypeSpelling Spelling
           deriving (Eq,Show)
-data Form = Form
-    { formCat :: Form_cat
-    } deriving (Eq,Show)
-data Form_cat = Form_cat_agglutination  |  Form_cat_register  | 
-                Form_cat_diacritic  |  Form_cat_homonymy  |  Form_cat_speling
-              deriving (Eq,Show)
-data Grammar = Grammar
-    { grammarCat :: Grammar_cat
-    } deriving (Eq,Show)
-data Grammar_cat = Grammar_cat_class  |  Grammar_cat_aux  | 
-                   Grammar_cat_gender  |  Grammar_cat_mode  |  Grammar_cat_number  | 
-                   Grammar_cat_person  |  Grammar_cat_tense  |  Grammar_cat_voice  | 
-                   Grammar_cat_euphony
-                 deriving (Eq,Show)
+data Form = FormLexical Lexical
+          | FormUnknown Unknown
+          deriving (Eq,Show)
+data Grammar = GrammarAgreement Agreement
+             | GrammarOmission Omission
+             | GrammarUnknown Unknown
+             deriving (Eq,Show)
 data Spelling = Spelling 		deriving (Eq,Show)
+data Unknown = Unknown 		deriving (Eq,Show)
+data Lexical = Lexical 		deriving (Eq,Show)
+data Agreement = Agreement 		deriving (Eq,Show)
+data Omission = Omission
+    { omissionType :: Omission_type
+    } deriving (Eq,Show)
+data Omission_type = Omission_type_verb  |  Omission_type_subject
+                      |  Omission_type_dirobject  |  Omission_type_indobject  | 
+                     Omission_type_preosition  |  Omission_type_other
+                   deriving (Eq,Show)
 newtype Context = Context String 		deriving (Eq,Show)
 data Comment = Comment Comment_Attrs String
              deriving (Eq,Show)
@@ -146,90 +149,113 @@ instance XmlContent Type where
         }
 
 instance HTypeable Form where
-    toHType x = Defined "Form" [] []
+    toHType x = Defined "form" [] []
 instance XmlContent Form where
-    toContents as =
-        [CElem (Elem "Form" (toAttrs as) []) ()]
-    parseContents = do
-        { (Elem _ as []) <- element ["Form"]
-        ; return (fromAttrs as)
-        } `adjustErr` ("in <Form>, "++)
[…]

### 2009-03-30 src/Annotator/DTD.hs
Added some more granularity.

@@ -23,6 +23,7 @@ data Error = Error Errtoks Type Context (Maybe Target)
 data Type = TypeForm Form
           | TypeGrammar Grammar
           | TypeSpelling Spelling
+          | TypeGibberish Gibberish
           deriving (Eq,Show)
 data Form = FormLexical Lexical
           | FormUnknown Unknown
@@ -32,6 +33,7 @@ data Grammar = GrammarAgreement Agreement
              | GrammarUnknown Unknown
              deriving (Eq,Show)
 data Spelling = Spelling 		deriving (Eq,Show)
+data Gibberish = Gibberish 		deriving (Eq,Show)
 data Unknown = Unknown 		deriving (Eq,Show)
 data Lexical = Lexical 		deriving (Eq,Show)
 data Agreement = Agreement 		deriving (Eq,Show)
@@ -40,14 +42,10 @@ data Omission = Omission
     } deriving (Eq,Show)
 data Omission_type = Omission_type_verb  |  Omission_type_subject
                       |  Omission_type_dirobject  |  Omission_type_indobject  | 
-                     Omission_type_preosition  |  Omission_type_other
+                     Omission_type_preposition  |  Omission_type_other
                    deriving (Eq,Show)
 newtype Context = Context String 		deriving (Eq,Show)
-data Comment = Comment Comment_Attrs String
-             deriving (Eq,Show)
-data Comment_Attrs = Comment_Attrs
-    { commentAuthor :: String
-    } deriving (Eq,Show)
+newtype Comment = Comment String 		deriving (Eq,Show)
 newtype Target = Target String 		deriving (Eq,Show)
 data Token = Token Token_Attrs String
            deriving (Eq,Show)
@@ -139,12 +137,15 @@ instance XmlContent Type where
         [CElem (Elem "type" [] (toContents a) ) ()]
     toContents (TypeSpelling a) =
         [CElem (Elem "type" [] (toContents a) ) ()]
+    toContents (TypeGibberish a) =
+        [CElem (Elem "type" [] (toContents a) ) ()]
     parseContents = do 
         { e@(Elem _ [] _) <- element ["type"]
         ; interior e $ oneOf
             [ return (TypeForm) `apply` parseContents
             , return (TypeGrammar) `apply` parseContents
             , return (TypeSpelling) `apply` parseContents
+            , return (TypeGibberish) `apply` parseContents
             ] `adjustErr` ("in <type>, "++)
         }
 
[…]

### 2009-03-30 src/Annotator/DTD.hs
Removed unnecessary context item.

@@ -17,8 +17,7 @@ data Tokens_Attrs = Tokens_Attrs
 data Errtoks = Errtoks
     { errtoksIdx :: String
     } deriving (Eq,Show)
-data Error = Error Errtoks Type Context (Maybe Target)
-                   (Maybe Comment)
+data Error = Error Errtoks Type (Maybe Target) (Maybe Comment)
            deriving (Eq,Show)
 data Type = TypeForm Form
           | TypeGrammar Grammar
@@ -44,7 +43,6 @@ data Omission_type = Omission_type_verb  |  Omission_type_subject
                       |  Omission_type_dirobject  |  Omission_type_indobject  | 
                      Omission_type_preposition  |  Omission_type_other
                    deriving (Eq,Show)
-newtype Context = Context String 		deriving (Eq,Show)
 newtype Comment = Comment String 		deriving (Eq,Show)
 newtype Target = Target String 		deriving (Eq,Show)
 data Token = Token Token_Attrs String
@@ -117,15 +115,14 @@ instance XmlAttributes Errtoks where
 instance HTypeable Error where
     toHType x = Defined "error" [] []
 instance XmlContent Error where
-    toContents (Error a b c d e) =
+    toContents (Error a b c d) =
         [CElem (Elem "error" [] (toContents a ++ toContents b ++
-                                 toContents c ++ maybe [] toContents d ++
-                                 maybe [] toContents e)) ()]
+                                 maybe [] toContents c ++ maybe [] toContents d)) ()]
     parseContents = do
         { e@(Elem _ [] _) <- element ["error"]
         ; interior e $ return (Error) `apply` parseContents
-                       `apply` parseContents `apply` parseContents
-                       `apply` optional parseContents `apply` optional parseContents
+                       `apply` parseContents `apply` optional parseContents
+                       `apply` optional parseContents
         } `adjustErr` ("in <error>, "++)
 
 instance HTypeable Type where
@@ -267,16 +264,6 @@ instance XmlAttrType Omission_type where
     toAttrFrTyp n Omission_type_indobject = Just (n, str2attr "indobject")
     toAttrFrTyp n Omission_type_preposition = Just (n, str2attr "preposition")
     toAttrFrTyp n Omission_type_other = Just (n, str2attr "other")
-
-instance HTypeable Context where
-    toHType x = Defined "context" [] []
-instance XmlContent Context where
-    toContents (Context a) =
[…]

### 2009-03-30 src/Annotator/Interface.hs
Added TreeViews and FormError handling.

@@ -17,6 +17,7 @@ module Annotator.Interface
 
 import Annotator.DTD
 import Annotator.Interface.Constants
+import Annotator.Interface.Models
 import Annotator.Interface.Types
 import Control.Monad.Trans (liftIO)
 import Data.IORef
@@ -28,7 +29,7 @@ import Graphics.UI.Gtk.Gdk.EventM
 import Graphics.UI.Gtk.Glade
 import Graphics.UI.Gtk.Windows.Dialog
 import Text.XML.HaXml.XmlContent.Haskell (readXml)
-import Text.XML.HaXml.XmlContent (fWriteXml)
+import Text.XML.HaXml.XmlContent (XmlContent, fWriteXml)
 
 -- Takes a corpus and returns a string representing the corpus' text in plain text.
 xmlToTokenString :: Corpus -> (Int,String,TokenMap)
@@ -143,7 +144,31 @@ initControls gui = do quitItem <- xmlGetWidget (xml gui) castToMenuItem menuItem
                       openItem `afterActivateLeaf` openItemHandler gui
                       clearBtn `onClicked` clearBtnHandler gui
                       spellBtn `onClicked` spellBtnHandler gui
+                      grmview  <- xmlGetWidget (xml gui) castToTreeView "grammarView"
+                      frmview  <- xmlGetWidget (xml gui) castToTreeView "formView"
+                      (initTreeView grmview) =<< grammarStore
+                      (initTreeView frmview) =<< formStore
+                      frecBtn <- xmlGetWidget (xml gui) castToButton "formRecordButton"
+                      grecBtn <- xmlGetWidget (xml gui) castToButton "grammarRecordButton"
+                      frecBtn `onClicked` frecBtnHandler gui frmview
+                      grecBtn `onClicked` grecBtnHandler gui grmview
+
                       return ()
+
+
+grecBtnHandler :: Gui -> TreeView -> IO ()
+grecBtnHandler gui view = undefined
+
+initTreeView :: (XmlContent a) => TreeView -> TreeStore (EType a) -> IO ()
+initTreeView view model =  do
+                        view `treeViewSetModel` model
+                        view `treeViewSetHeadersVisible` False
+                        column <- treeViewColumnNew
+                        renderer <- cellRendererTextNew
+                        cellLayoutPackStart column renderer True
+                        cellLayoutSetAttributes column renderer model $ \row -> [cellText := name row]
+                        view `treeViewAppendColumn` column
+                        return ()
 
 openItemHandler :: Gui -> IO ()
[…]

### 2009-03-30 src/Annotator/Interface/Models.hs

@@ -1,0 +1,43 @@
+module Annotator.Interface.Models where
+    
+import qualified Data.Tree as Tree
+import qualified Data.List as List
+
+import Annotator.DTD
+import Graphics.UI.Gtk
+import Graphics.UI.Gtk.Glade
+import Graphics.UI.Gtk.ModelView
+import Text.XML.HaXml.XmlContent
+
+data (XmlContent a) => EType a =  EType { name :: String , xmlContent :: Maybe a }
+
+grammarStore :: IO (TreeStore (EType Grammar))
+grammarStore = treeStoreNew
+        [ Tree.Node
+           { Tree.rootLabel = EType { name = "Omission", xmlContent = Nothing }
+           , Tree.subForest = [ leafNode EType { name = "subject" 
+                                               , xmlContent = Just $ GrammarOmission (Omission Omission_type_subject)}
+                              , leafNode EType { name = "verb" 
+                                               , xmlContent = Just $ GrammarOmission (Omission Omission_type_verb)}
+                              , leafNode EType { name = "direct object" 
+                                               , xmlContent = Just $ GrammarOmission (Omission Omission_type_dirobject)}
+                              , leafNode EType { name = "indirect object" 
+                                               , xmlContent = Just $ GrammarOmission (Omission Omission_type_indobject)}
+                              , leafNode EType { name = "preposition" 
+                                               , xmlContent = Just $ GrammarOmission (Omission Omission_type_preposition)}
+                              , leafNode EType { name = "other" 
+                                               , xmlContent = Just $ GrammarOmission (Omission Omission_type_other)}
+                              ]
+           }
+        , leafNode EType { name = "Agreement", xmlContent = Just $ GrammarAgreement Agreement }
+        , leafNode EType { name = "Unknown", xmlContent = Just $ GrammarUnknown Unknown }
+        ]
+
+formStore = treeStoreNew
+        [ leafNode EType { name = "Lexical", xmlContent = Just $ FormLexical Lexical }
+        , leafNode EType { name = "Unknown", xmlContent = Just $ FormUnknown Unknown }
+        ]
+
+leafNode :: a -> Tree.Tree a
+leafNode a = Tree.Node { Tree.rootLabel = a, Tree.subForest = [] }
+

### 2009-03-30 src/Annotator/Interface.hs
Minor refactoring.

@@ -155,7 +155,6 @@ initControls gui = do quitItem <- xmlGetWidget (xml gui) castToMenuItem menuItem
 
                       return ()
 
-
 grecBtnHandler :: Gui -> TreeView -> IO ()
 grecBtnHandler gui view = undefined
 
@@ -205,8 +204,7 @@ frecBtnHandler gui view = do row <- (treeViewGetSelection view >>= treeSelection
                                   Nothing -> showError "Please select some tokens first"
                                   Just [] -> showError "Please select some tokens first"
                                   Just ts -> case row of
-                                                  [path] -> do store <- formStore
-                                                               (EType _ content) <- store `treeStoreGetValue` path
+                                                  [path] -> do (EType _ content) <- (\s -> s `treeStoreGetValue` path) =<< formStore
                                                                case content of
                                                                     Nothing -> showError "Select a leaf."
                                                                     Just etype -> do addToErrors gui e

### 2009-03-31 src/Annotator/Interface.hs
Some major refactoring and trigger Button.

@@ -62,34 +62,32 @@ xmlFileFilter = do ff <- fileFilterNew
                    fileFilterAddMimeType ff "text/xml"
                    return ff
 
-findContext :: Gui -> EventM EButton Bool
-findContext gui =
+findContext :: Gui -> Label -> IORef [Token] -> EventM EButton Bool
+findContext gui l ref =
     do btn <- eventButton
        case btn of
            LeftButton -> do 
                coords <- eventCoordinates
                let (x,y) = truncCoordToInt coords
-               liftIO $ do findToken gui (x,y)
+               liftIO $ do findToken gui (x,y) l ref 
                            return False
            _          -> return False
        
 truncCoordToInt :: (Double,Double) -> (Int,Int)
 truncCoordToInt (x,y) = (truncate x,truncate y)
 
-findToken :: Gui -> (Int,Int) -> IO ()
-findToken gui (x,y) = do
+findToken :: Gui -> (Int,Int) -> Label -> IORef [Token] -> IO ()
+findToken gui (x,y) l tokenRef = do
         bcrd <- (corpusView gui) `textViewWindowToBufferCoords` TextWindowWidget $ (x,y)
         iter <- uncurry (textViewGetIterAtLocation (corpusView gui)) bcrd
         loc <- textIterGetOffset iter
         Just tmap <- readIORef (tokens gui)
         let Just token = (Point loc) `M.lookup` tmap
-        ref <- readIORef (selectedTkn gui)
-        case ref of
-             Nothing -> writeIORef (selectedTkn gui) (Just [token])
-             Just toks -> if (not (elem token toks))
-                               then writeIORef (selectedTkn gui) (Just $ toks ++ [token])
-                               else writeIORef (selectedTkn gui) (Just $ delete token toks)
-        putTokensOnLabel gui
+        toks <- readIORef tokenRef
+        if (not (elem token toks))
+           then writeIORef tokenRef $ token:toks
+           else writeIORef tokenRef $ delete token toks
+        putTokensOnLabel gui l
                         
 -- | Entry to the GUI
 runGUI :: IO ()
@@ -118,8 +116,9 @@ prepareGUI = do
                              textView  <- xmlGetWidget gladeXml castToTextView "corpusView"
                              nothingRef <- newIORef Nothing
                              nothingRef' <- newIORef Nothing
-                             nothingRef'' <- newIORef Nothing
+                             nothingRef'' <- newIORef []
                              nothingRef''' <- newIORef Nothing
[…]

### 2009-03-31 src/Annotator/Interface/Types.hs

@@ -17,7 +17,8 @@ data Gui = Gui { corpusView  :: TextView -- ^ The corpusView
                , tokenLabel  :: Label -- ^ The Label displaying tokens
                , corpusClick :: IORef (Maybe (ConnectId TextView))
                , tokens      :: IORef (Maybe TokenMap)
-               , selectedTkn :: IORef (Maybe [Token])
+               , selectedTkn :: IORef [Token]
+               , trigger     :: IORef [Token]
                , xmlDocument :: IORef (Maybe Corpus)
                }
 

### 2009-04-05 src/Annotator/Interface.hs
Added error view.

@@ -62,6 +62,7 @@ prepareGUI = do
                              nothingRef'' <- newIORef []
                              nothingRef''' <- newIORef Nothing
                              nothingRef'''' <- newIORef []
+                             listRef <- newIORef =<< (listStoreNew [])
                              somethingRef <- newIORef nothingRef''
                              let gui = Gui { corpusView  = textView
                                            , window      = w
@@ -71,6 +72,7 @@ prepareGUI = do
                                            , tokenArray  = nothingRef'
                                            , selectedTkn = nothingRef''
                                            , xmlDocument = nothingRef'''
+                                           , errorModel  = listRef
                                            , currentFocs = somethingRef
                                            , trigger     = nothingRef''''
                                            }
@@ -108,7 +110,9 @@ initControls gui = do quitItem <- xmlGetWidget (xml gui) castToMenuItem menuItem
                       openItem `afterActivateLeaf` openItemHandler gui
                       clearBtn `onClicked` clearBtnHandler gui
                       errorView <- xmlGetWidget (xml gui) castToTreeView "errorView"
+                      listView <- xmlGetWidget (xml gui) castToTreeView "treeView1"
                       (initTreeView errorView) =<< errorStore
+                      (initListView listView) =<< (readIORef (errorModel gui))
                       recordBtn <- xmlGetWidget (xml gui) castToButton "recordButton"
                       recordBtn `onClicked` recordHandler gui errorView
                       (triggerBtn gui) `afterToggled` triggerToggleHandler gui
@@ -124,6 +128,38 @@ initTreeView view model =  do
                         cellLayoutSetAttributes column renderer model $ \row -> [cellText := name row]
                         view `treeViewAppendColumn` column
                         return ()
+
+initListView :: TreeView -> ListStore Record -> IO ()
+initListView view model = do
+    view `treeViewSetModel` model
+    view `treeViewSetHeadersVisible` True
+    
+    columnTokens <- treeViewColumnNew
+    rendererTokens <- cellRendererTextNew
+    treeViewColumnPackStart columnTokens rendererTokens True
+    cellLayoutSetAttributes columnTokens rendererTokens model $ \row -> [cellText := record2tokens row ]
+    treeViewColumnSetTitle columnTokens "Error Tokens"
[…]

### 2009-04-05 src/Annotator/Interface/Handlers.hs

@@ -27,7 +27,9 @@ addToRecords :: Gui -> Record -> IO ()
 addToRecords gui err = do ref <- readIORef (xmlDocument gui)
                           case ref of
                               Nothing  -> putStrLn "Warning, empty document!"
-                              Just doc -> writeIORef (xmlDocument gui) (Just $ ate err doc) 
+                              Just doc -> do writeIORef (xmlDocument gui) (Just $ ate err doc) 
+                                             emodel <- readIORef (errorModel gui)
+                                             emodel `listStorePrepend` err
                               where ate e c@(Corpus ts (Errors es)) = 
                                                              if e `elem` es
                                                                 then c

### 2009-04-05 src/Annotator/Interface/Types.hs

@@ -19,6 +19,7 @@ data Gui = Gui { corpusView  :: TextView -- ^ The corpusView
                , triggerBtn  :: ToggleButton
                , selectedTkn :: IORef [Token]
                , trigger     :: IORef [Token]
+               , errorModel  :: IORef (ListStore Record)
                , currentFocs :: IORef (IORef [Token])
                , tokenArray  :: IORef (Maybe (Array Int Token))
                , xmlDocument :: IORef (Maybe Corpus)

### 2009-04-05 src/Annotator/DTD.hs
Revised and finished taxonomy.

@@ -35,9 +35,13 @@ data Error_ = Error_Context Context
 newtype Context = Context (Maybe Context_) 		deriving (Eq,Show)
 data Context_ = Context_Lexical Lexical
               | Context_Idiom Idiom
+              | Context_Verbtense Verbtense
               deriving (Eq,Show)
 newtype Grammar = Grammar (Maybe Grammar_) 		deriving (Eq,Show)
 data Grammar_ = Grammar_Agreement Agreement
+              | Grammar_Voice Voice
+              | Grammar_Valency Valency
+              | Grammar_Complementation Complementation
               | Grammar_Omission Omission
               | Grammar_Wo Wo
               | Grammar_Morphology Morphology
@@ -74,6 +78,10 @@ data Morphology_ = Morphology_Derivation Derivation
                  | Morphology_Compounding Compounding
                  | Morphology_Inflection Inflection
                  deriving (Eq,Show)
+newtype Wo = Wo (Maybe Wo_) 		deriving (Eq,Show)
+data Wo_ = Wo_Internal Internal
+         | Wo_External External
+         deriving (Eq,Show)
 data Number = Number 		deriving (Eq,Show)
 data Gender = Gender 		deriving (Eq,Show)
 data Case = Case 		deriving (Eq,Show)
@@ -83,14 +91,19 @@ data Subject = Subject 		deriving (Eq,Show)
 data Predicate = Predicate 		deriving (Eq,Show)
 data Article = Article 		deriving (Eq,Show)
 data Preposition = Preposition 		deriving (Eq,Show)
-data Wo = Wo 		deriving (Eq,Show)
 data Redundancy = Redundancy 		deriving (Eq,Show)
 data Homonymy = Homonymy 		deriving (Eq,Show)
 data Register = Register 		deriving (Eq,Show)
+data External = External 		deriving (Eq,Show)
+data Internal = Internal 		deriving (Eq,Show)
 data Diacritics = Diacritics 		deriving (Eq,Show)
+data Verbtense = Verbtense 		deriving (Eq,Show)
 data Local = Local 		deriving (Eq,Show)
 data Lexis = Lexis 		deriving (Eq,Show)
 data Idiom = Idiom 		deriving (Eq,Show)
+data Voice = Voice 		deriving (Eq,Show)
+data Valency = Valency 		deriving (Eq,Show)
+data Complementation = Complementation 		deriving (Eq,Show)
[…]

### 2009-04-05 src/Annotator/Interface/Models.hs

@@ -24,7 +24,38 @@ errorStore = treeStoreNew
         , leafNode $ EType "Diacritics" (Error  (Just $ Error_Spelling (Spelling (Just $ Spelling_Diacritics Diacritics))))
         , leafNode $ EType "Local" (Error  (Just $ Error_Spelling (Spelling (Just $ Spelling_Local Local))))
         ]
-      , leafNode $ EType "Grammar" (Error  (Just $ Error_Grammar (Grammar  Nothing)))
+      , Tree.Node (EType "Grammar" (Error  (Just $ Error_Grammar (Grammar  Nothing))))
+        [ Tree.Node (EType "Agreement" (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Agreement (Agreement Nothing))))))
+          [ leafNode $ EType "Number" (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Agreement (Agreement (Just $ Agreement_Number Number))))))
+          , leafNode $ EType "Gender" (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Agreement (Agreement (Just $ Agreement_Gender Gender))))))
+          , leafNode $ EType "Case" (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Agreement (Agreement (Just $ Agreement_Case Case))))))
+          , leafNode $ EType "Person" (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Agreement (Agreement (Just $ Agreement_Person Person))))))
+          , leafNode $ EType "Tense" (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Agreement (Agreement (Just $ Agreement_Tense Tense))))))
+          ]
+        , leafNode $ EType "Redundancy" (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Redundancy Redundancy))))
+        , leafNode $ EType "Valency" (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Valency Valency))))
+        , leafNode $ EType "Complementation" (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Complementation Complementation))))
+        , leafNode $ EType "Voice" (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Voice Voice))))
+        , Tree.Node (EType "Word Order" (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Wo (Wo Nothing))))))
[…]

### 2009-04-05 src/Annotator/Interface.hs
Added laoding of new errors on reading file.

@@ -21,7 +21,7 @@ import Annotator.Interface.Models
 import Annotator.Interface.Handlers
 import Annotator.Interface.Util
 import Annotator.Interface.Types
-import Control.Monad (forM)
+import Control.Monad (forM,forM_)
 import Data.IORef
 import GHC.List hiding (span)
 import Graphics.UI.Gtk
@@ -202,7 +202,7 @@ loadFile gui fn = do
                readCorpus c tb gui
 
 readCorpus :: Corpus -> TextBuffer -> Gui -> IO ()
-readCorpus corpus@(Corpus (Tokens _ _) (Errors _)) tb gui =
+readCorpus corpus@(Corpus (Tokens _ _) (Errors es)) tb gui =
         do putStrLn "Indexing tokens..."
            updateRef (tokenArray gui) tokens
            putStrLn "Filling Buffer..."
@@ -212,6 +212,8 @@ readCorpus corpus@(Corpus (Tokens _ _) (Errors _)) tb gui =
            tags <- forM tokenList (token2Tag gui tt)
            putStrLn "Applying tags..."
            applyTags tb tags tokenList
+           putStrLn "Reading in records..."
+           forM_ es (addToErrorView gui)
            putStrLn "Finished."
            where tokens = xmlToArray corpus
                  tokenList = elems tokens

### 2009-04-05 src/Annotator/Interface/Handlers.hs

@@ -28,13 +28,16 @@ addToRecords gui err = do ref <- readIORef (xmlDocument gui)
                           case ref of
                               Nothing  -> putStrLn "Warning, empty document!"
                               Just doc -> do writeIORef (xmlDocument gui) (Just $ ate err doc) 
-                                             emodel <- readIORef (errorModel gui)
-                                             emodel `listStorePrepend` err
+                                             addToErrorView gui err
                               where ate e c@(Corpus ts (Errors es)) = 
                                                              if e `elem` es
                                                                 then c
                                                                 else Corpus ts (Errors (e:es))
 
+addToErrorView :: Gui -> Record -> IO ()
+addToErrorView gui record =  
+        (\model -> model `listStorePrepend` record) =<< (readIORef (errorModel gui))
+                                             
 makeRecord :: Gui -> Error -> IO Record
 makeRecord gui etype = do triggers <- readIORef (trigger gui)
                           tokens  <- readIORef (selectedTkn gui)

### 2009-04-05 src/Annotator/Interface/Handlers.hs
Fixed clear button bug:
Wouldn't reset toggleButton's state (but would function correct still.
Minor nuisance).

@@ -50,6 +50,7 @@ clearBtnHandler :: Gui -> IO ()
 clearBtnHandler gui =  do writeIORef (selectedTkn gui) []
                           writeIORef (trigger     gui) []
                           writeIORef (currentFocs gui) (selectedTkn gui)
+                          (triggerBtn gui) `toggleButtonSetActive` False
                           putTokensOnLabels gui
 
 tagEventHandler :: Gui -> Token -> Old.Event -> TextIter -> IO ()

### 2009-04-05 src/Annotator/Interface.hs
Added delete button functionality.

@@ -109,11 +109,13 @@ initControls gui = do quitItem <- xmlGetWidget (xml gui) castToMenuItem menuItem
                       quitItem `afterActivateLeaf` widgetDestroy (window gui)
                       openItem `afterActivateLeaf` openItemHandler gui
                       clearBtn `onClicked` clearBtnHandler gui
+                      deleteBtn <- xmlGetWidget (xml gui) castToButton "deleteButton"
                       errorView <- xmlGetWidget (xml gui) castToTreeView "errorView"
                       listView <- xmlGetWidget (xml gui) castToTreeView "treeView1"
                       (initTreeView errorView) =<< errorStore
                       (initListView listView) =<< (readIORef (errorModel gui))
                       recordBtn <- xmlGetWidget (xml gui) castToButton "recordButton"
+                      deleteBtn `onClicked` deleteHandler gui listView
                       recordBtn `onClicked` recordHandler gui errorView
                       (triggerBtn gui) `afterToggled` triggerToggleHandler gui
                       return ()
@@ -178,9 +180,10 @@ openItemHandler gui = do fn <- openFileAction gui
                               _          -> return ()
 
 saveItemHandler :: Gui -> FilePath -> IO ()
-saveItemHandler gui fn = do ref <- readIORef (xmlDocument gui)
+saveItemHandler gui fn = do errors <- listStoreToList =<< (readIORef $ errorModel gui)
+                            ref <- readIORef (xmlDocument gui)
                             case ref of
-                                 Just doc -> fWriteXml fn doc
+                                 Just (Corpus ts _) -> fWriteXml fn (Corpus ts (Errors errors))
                                  Nothing  -> showError "Load a file first"
 
 saveAsItemHandler :: Gui -> IO ()

### 2009-04-05 src/Annotator/Interface/Handlers.hs

@@ -4,11 +4,22 @@ import Annotator.Interface.Types
 import Annotator.Interface.Util
 import Annotator.Interface.Models
 import Annotator.DTD
+import Control.Monad (forM_)
 import Data.IORef
 import Data.List
 import GHC.List hiding (span)
 import Graphics.UI.Gtk
 import qualified Graphics.UI.Gtk.Gdk.Events as Old
+
+deleteHandler :: Gui -> TreeView -> IO ()
+deleteHandler gui view = do
+    rows <- treeSelectionGetSelectedRows =<< (treeViewGetSelection view)
+    forM_ rows (deleteThisRecord gui)
+
+deleteThisRecord :: Gui -> TreePath -> IO ()
+deleteThisRecord gui [index] = do model <- readIORef (errorModel gui)
+                                  model `listStoreRemove` index
+deleteThisRecord _ _ = putStrLn "Panic! We got a tree path and wanted a list index"
 
 recordHandler :: Gui -> TreeView -> IO ()
 recordHandler gui view = 
@@ -19,21 +30,9 @@ recordHandler gui view =
                  case paths of
                    [path] -> do (EType _ etype) <- (\store ->treeStoreGetValue store path) =<< errorStore
                                 record <- makeRecord gui etype
-                                addToRecords gui record
-                                clearBtnHandler gui
+                                addToErrorView gui record
                    _ -> putStrLn "Warning: Select one error type first."
        
-addToRecords :: Gui -> Record -> IO ()
-addToRecords gui err = do ref <- readIORef (xmlDocument gui)
-                          case ref of
-                              Nothing  -> putStrLn "Warning, empty document!"
-                              Just doc -> do writeIORef (xmlDocument gui) (Just $ ate err doc) 
-                                             addToErrorView gui err
-                              where ate e c@(Corpus ts (Errors es)) = 
-                                                             if e `elem` es
-                                                                then c
-                                                                else Corpus ts (Errors (e:es))
-
 addToErrorView :: Gui -> Record -> IO ()
 addToErrorView gui record =  
         (\model -> model `listStorePrepend` record) =<< (readIORef (errorModel gui))

### 2009-04-05 src/Annotator/Interface.hs
Some work on displaying error tokens correctly.
I'm getting some weird array errors here…

@@ -58,7 +58,7 @@ prepareGUI = do
                              tl       <- xmlGetWidget gladeXml castToLabel "tokenLabel"
                              textView  <- xmlGetWidget gladeXml castToTextView "corpusView"
                              trigButton  <- xmlGetWidget gladeXml castToToggleButton "triggerButton"
-                             nothingRef' <- newIORef Nothing
+                             nothingRef' <- newIORef (array (0,0) [])
                              nothingRef'' <- newIORef []
                              nothingRef''' <- newIORef Nothing
                              nothingRef'''' <- newIORef []
@@ -113,7 +113,7 @@ initControls gui = do quitItem <- xmlGetWidget (xml gui) castToMenuItem menuItem
                       errorView <- xmlGetWidget (xml gui) castToTreeView "errorView"
                       listView <- xmlGetWidget (xml gui) castToTreeView "treeView1"
                       (initTreeView errorView) =<< errorStore
-                      (initListView listView) =<< (readIORef (errorModel gui))
+                      (initListView gui listView) =<< (readIORef (errorModel gui))
                       recordBtn <- xmlGetWidget (xml gui) castToButton "recordButton"
                       deleteBtn `onClicked` deleteHandler gui listView
                       recordBtn `onClicked` recordHandler gui errorView
@@ -131,22 +131,24 @@ initTreeView view model =  do
                         view `treeViewAppendColumn` column
                         return ()
 
-initListView :: TreeView -> ListStore Record -> IO ()
-initListView view model = do
+initListView :: Gui -> TreeView -> ListStore Record -> IO ()
+initListView gui view model = do
     view `treeViewSetModel` model
     view `treeViewSetHeadersVisible` True
+    
+    tokens <- readIORef (tokenArray gui)
     
     columnTokens <- treeViewColumnNew
     rendererTokens <- cellRendererTextNew
     treeViewColumnPackStart columnTokens rendererTokens True
-    cellLayoutSetAttributes columnTokens rendererTokens model $ \row -> [cellText := record2tokens row ]
+    cellLayoutSetAttributes columnTokens rendererTokens model $ \row -> [cellText := record2tokens row tokens ]
     treeViewColumnSetTitle columnTokens "Error Tokens"
[…]

### 2009-04-05 src/Annotator/Interface/Types.hs

@@ -21,7 +21,7 @@ data Gui = Gui { corpusView  :: TextView -- ^ The corpusView
                , trigger     :: IORef [Token]
                , errorModel  :: IORef (ListStore Record)
                , currentFocs :: IORef (IORef [Token])
-               , tokenArray  :: IORef (Maybe (Array Int Token))
+               , tokenArray  :: IORef (Array Int Token)
                , xmlDocument :: IORef (Maybe Corpus)
                }
 

### 2009-04-05 src/Annotator/Interface/Handlers.hs
Fixed record button bug.

@@ -31,6 +31,7 @@ recordHandler gui view =
                    [path] -> do (EType _ etype) <- (\store ->treeStoreGetValue store path) =<< errorStore
                                 record <- makeRecord gui etype
                                 addToErrorView gui record
+                                clearBtnHandler gui
                    _ -> putStrLn "Warning: Select one error type first."
        
 addToErrorView :: Gui -> Record -> IO ()

### 2009-04-05 src/Annotator/Interface.hs
Enhanced usability: highlight, text tokens, index.

@@ -112,8 +112,8 @@ initControls gui = do quitItem <- xmlGetWidget (xml gui) castToMenuItem menuItem
                       deleteBtn <- xmlGetWidget (xml gui) castToButton "deleteButton"
                       errorView <- xmlGetWidget (xml gui) castToTreeView "errorView"
                       listView <- xmlGetWidget (xml gui) castToTreeView "treeView1"
+                      tokens <- readIORef (tokenArray gui)
                       (initTreeView errorView) =<< errorStore
-                      (initListView gui listView) =<< (readIORef (errorModel gui))
                       recordBtn <- xmlGetWidget (xml gui) castToButton "recordButton"
                       deleteBtn `onClicked` deleteHandler gui listView
                       recordBtn `onClicked` recordHandler gui errorView
@@ -131,12 +131,10 @@ initTreeView view model =  do
                         view `treeViewAppendColumn` column
                         return ()
 
-initListView :: Gui -> TreeView -> ListStore Record -> IO ()
-initListView gui view model = do
+initListView :: Array Int Token -> TreeView -> ListStore Record -> IO ()
+initListView tokens view model = do
     view `treeViewSetModel` model
     view `treeViewSetHeadersVisible` True
-    
-    tokens <- readIORef (tokenArray gui)
     
     columnTokens <- treeViewColumnNew
     rendererTokens <- cellRendererTextNew
@@ -152,23 +150,32 @@ initListView gui view model = do
     treeViewColumnSetTitle columnTrigger "Trigger"
     view `treeViewAppendColumn` columnTrigger
     
+    columnIndex <- treeViewColumnNew
+    rendererIndex <- cellRendererTextNew
+    treeViewColumnPackStart columnIndex rendererIndex True
+    cellLayoutSetAttributes columnIndex rendererIndex model $ \row -> [cellText := record2index row ]
+    treeViewColumnSetTitle columnIndex "Index"
+    view `treeViewAppendColumn` columnIndex
+
     columnType <- treeViewColumnNew
     rendererType <- cellRendererTextNew
     treeViewColumnPackStart columnType rendererType True
     cellLayoutSetAttributes columnType rendererType model $ \row -> [cellText := record2type row ]
     treeViewColumnSetTitle columnType "Type"
     view `treeViewAppendColumn` columnType
-    
+
     return ()
 
[…]

### 2009-04-05 src/Annotator/Interface/Handlers.hs

@@ -9,7 +9,9 @@ import Data.IORef
 import Data.List
 import GHC.List hiding (span)
 import Graphics.UI.Gtk
+import Graphics.UI.Gtk.Glade
 import qualified Graphics.UI.Gtk.Gdk.Events as Old
+import System.Glib.Attributes
 
 deleteHandler :: Gui -> TreeView -> IO ()
 deleteHandler gui view = do
@@ -33,15 +35,25 @@ recordHandler gui view =
                                 addToErrorView gui record
                                 clearBtnHandler gui
                    _ -> putStrLn "Warning: Select one error type first."
-       
+
 addToErrorView :: Gui -> Record -> IO ()
-addToErrorView gui record =  
+addToErrorView gui record@(Record _ (Errtoks tokens) _ _ _) = do
         (\model -> model `listStorePrepend` record) =<< (readIORef (errorModel gui))
-                                             
+        highlightTags gui (words tokens)
+
 makeRecord :: Gui -> Error -> IO Record
 makeRecord gui etype = do triggers <- readIORef (trigger gui)
                           tokens  <- readIORef (selectedTkn gui)
                           return $ Record (Record_Attrs (Just $ token2string triggers) Nothing) (Errtoks $ token2string tokens) etype Nothing Nothing
+
+highlightTags gui tokens = do
+        buffer <- textViewGetBuffer (corpusView gui)
+        table <- textBufferGetTagTable buffer
+        forM_ tokens (\idx -> do
+                mtag <- table `textTagTableLookup` idx
+                case mtag of
+                        Nothing -> putStrLn "Warning: token tag not in table. Bad."
+                        Just tag -> set tag [ textTagBackground := "#dd8888" ])
 
 token2string :: [Token] -> String
 token2string = unwords . (map tokenId)
@@ -54,6 +66,9 @@ clearBtnHandler gui =  do writeIORef (selectedTkn gui) []
                           putTokensOnLabels gui
 
 tagEventHandler :: Gui -> Token -> Old.Event -> TextIter -> IO ()
+tagEventHandler gui (Token (Token_Attrs t) _ ) (Old.Motion _ _ _ _ _ _ _ _ ) _ = do
+    label <- xmlGetWidget (xml gui) castToLabel "tindexLabel"
+    label `labelSetText` (drop 1 t)
 tagEventHandler gui t (Old.Button _ Old.SingleClick _ _ _ _ Old.LeftButton  _ _) _ = do
     tref <- readIORef (currentFocs gui)
     seltokens <- readIORef tref

### 2009-04-06 src/tools/Agreement.hs
Added agreement program stub.

@@ -1,0 +1,44 @@
+module Main where
+
+import Annotator.DTD
+import Control.Monad
+import Data.Either
+import Data.List
+import System.Environment
+
+import Text.XML.HaXml.XmlContent.Haskell (readXml)
+
+data ComparedRecords = ComparedRecords { agreement    :: [RecordPair]
+                                       , disagreement :: ([Record],[Record]) }
+data RecordPair = RecordPair Record Record
+    
+main :: IO ()
+main = do args <- getArgs
+          case args of
+              
+              files@(_:_:_) -> do parses <- forM files (\f -> fmap readXml (readFile f))
+                                  case lefts parses of
+                                       [] -> compareCorpora $ rights parses
+                                       errors -> do forM_ errors putStrLn
+              _       -> putStrLn "Usage: agreement a1.xml a2.xml"
+
+compareCorpora :: [Corpus] -> IO ()
+compareCorpora ((Corpus (Tokens a1 ts1) (Errors es1)):(Corpus (Tokens a2 ts2) (Errors es2)):[]) = do
+    putStrLn "Testing for token to token Consistency..."
+    if a1 /= a2 || ts1 /= ts2
+       then putStrLn "Tokens inconsistent. Checking agreement doesn't make sense this way."
+       else do putStrLn "Tokens consistent."
+               let results = compareErrors2 es1 es2
+               undefined
+compareCorpora _          = putStrLn "Sorry, but only two corpora can be evaluated at this point."
+
+compareErrors2 :: [Record] -> [Record] -> ComparedRecords
+compareErrors2 es1 es2 = let ses1 = sortBy tokenEquality es1
+                             ses2 = sortBy tokenEquality es2
+                         in trawlErrors ses1 ses2 (ComparedRecords [] ([],[]))
+
+trawlErrors :: [Record] -> [Record] -> ComparedRecords -> ComparedRecords
+trawlErrors = undefined
+
+tokenEquality :: Record -> Record -> Ordering
+tokenEquality (Record _ (Errtoks t1) _ _ _) (Record _ (Errtoks t2) _ _ _) = compare t1 t2

### 2009-04-06 src/tools/Agreement.hs
Defined trawlErrors

@@ -1,6 +1,7 @@
 module Main where
 
 import Annotator.DTD
+import Annotator.Interface.Util
 import Control.Monad
 import Data.Either
 import Data.List
@@ -11,11 +12,11 @@ import Text.XML.HaXml.XmlContent.Haskell (readXml)
 data ComparedRecords = ComparedRecords { agreement    :: [RecordPair]
                                        , disagreement :: ([Record],[Record]) }
 data RecordPair = RecordPair Record Record
-    
+
 main :: IO ()
 main = do args <- getArgs
           case args of
-              
+
               files@(_:_:_) -> do parses <- forM files (\f -> fmap readXml (readFile f))
                                   case lefts parses of
                                        [] -> compareCorpora $ rights parses
@@ -38,7 +39,15 @@ compareErrors2 es1 es2 = let ses1 = sortBy tokenEquality es1
                          in trawlErrors ses1 ses2 (ComparedRecords [] ([],[]))
 
 trawlErrors :: [Record] -> [Record] -> ComparedRecords -> ComparedRecords
-trawlErrors = undefined
+trawlErrors [] _ c = c
+trawlErrors _ [] c = c
+trawlErrors (r1@(Record _ (Errtoks t1) _ _ _):rs1)
+            (r2@(Record _ (Errtoks t2) _ _ _):rs2)
+            (ComparedRecords agree disagree@(lrds,rrds))
+            | t1 == t2  = trawlErrors rs1 rs2
+                                      (ComparedRecords ((RecordPair r1 r2):agree) disagree)
+            | otherwise = trawlErrors rs1 rs2
+                                      (ComparedRecords agree ((r1:lrds),(r2:rrds)))
 
 tokenEquality :: Record -> Record -> Ordering
 tokenEquality (Record _ (Errtoks t1) _ _ _) (Record _ (Errtoks t2) _ _ _) = compare t1 t2

### 2009-04-06 src/Annotator/Interface.hs
Improved rendering of error type.

@@ -135,21 +135,21 @@ initListView :: Array Int Token -> TreeView -> ListStore Record -> IO ()
 initListView tokens view model = do
     view `treeViewSetModel` model
     view `treeViewSetHeadersVisible` True
-    
+
     columnTokens <- treeViewColumnNew
     rendererTokens <- cellRendererTextNew
     treeViewColumnPackStart columnTokens rendererTokens True
     cellLayoutSetAttributes columnTokens rendererTokens model $ \row -> [cellText := record2tokens row tokens ]
     treeViewColumnSetTitle columnTokens "Error Tokens"
     view `treeViewAppendColumn` columnTokens
-    
+
     columnTrigger <- treeViewColumnNew
     rendererTrigger <- cellRendererTextNew
     treeViewColumnPackStart columnTrigger rendererTrigger True
     cellLayoutSetAttributes columnTrigger rendererTrigger model $ \row -> [cellText := record2trigger row tokens ]
     treeViewColumnSetTitle columnTrigger "Trigger"
     view `treeViewAppendColumn` columnTrigger
-    
+
     columnIndex <- treeViewColumnNew
     rendererIndex <- cellRendererTextNew
     treeViewColumnPackStart columnIndex rendererIndex True
@@ -166,13 +166,13 @@ initListView tokens view model = do
 
     return ()
 
-record2tokens (Record _ (Errtoks errtoks) _ _ _) tokens = 
+record2tokens (Record _ (Errtoks errtoks) _ _ _) tokens =
         show $ map (tokenString . (tokens!) . read . drop 1) (words errtoks)
 record2index (Record _ (Errtoks errtoks) _ _ _) = drop 1 $ head $ words errtoks
 record2trigger (Record (Record_Attrs Nothing _) _ _ _ _) tokens = "N/A"
 record2trigger (Record (Record_Attrs (Just context) _) _ _ _ _) tokens = 
         show $ map (tokenString . (tokens!) . read . drop 1) (words context)
-record2type (Record _ _ etype _ _) = show etype
+record2type (Record _ _ etype _ _) = show $ shortenErrorType etype
 
 lookupIndices array = map (tokenString . (array!))
 

### 2009-04-06 src/tools/Agreement.hs
Added some instances.

@@ -10,8 +10,9 @@ import System.Environment
 import Text.XML.HaXml.XmlContent.Haskell (readXml)
 
 data ComparedRecords = ComparedRecords { agreement    :: [RecordPair]
-                                       , disagreement :: ([Record],[Record]) }
-data RecordPair = RecordPair Record Record
+                                       , disagreement :: ([Record],[Record]) } 
+                                       deriving Show, Eq
+data RecordPair = RecordPair Record Record deriving Eq
 
 main :: IO ()
 main = do args <- getArgs

### 2009-04-06 src/Annotator/Interface.hs
Added Comment and Target functionality.

@@ -22,9 +22,11 @@ import Annotator.Interface.Handlers
 import Annotator.Interface.Util
 import Annotator.Interface.Types
 import Control.Monad (forM,forM_)
+import Control.Monad.Trans (liftIO)
 import Data.IORef
 import GHC.List hiding (span)
 import Graphics.UI.Gtk
+import Graphics.UI.Gtk.Gdk.EventM
 import Graphics.UI.Gtk.Glade
 import Graphics.UI.Gtk.Windows.Dialog
 import Text.XML.HaXml.XmlContent.Haskell (readXml)
@@ -118,7 +120,26 @@ initControls gui = do quitItem <- xmlGetWidget (xml gui) castToMenuItem menuItem
                       deleteBtn `onClicked` deleteHandler gui listView
                       recordBtn `onClicked` recordHandler gui errorView
                       (triggerBtn gui) `afterToggled` triggerToggleHandler gui
+                      targetField <- xmlGetWidget (xml gui) castToEntry "targetField"
+                      commentField <- xmlGetWidget (xml gui) castToEntry "commentField"
+                      targetField `on` focusInEvent $ deleteField "Target Hypothesis" targetField
+                      commentField `on` focusInEvent $ deleteField "Comment" commentField
+                      targetField `on` focusOutEvent $ restoreField "Target Hypothesis" targetField
+                      commentField `on` focusOutEvent $ restoreField "Comment" commentField
                       return ()
+
+deleteField :: String -> Entry -> EventM EFocus Bool
+deleteField s e = liftIO $ do text <- entryGetText e
+                              if text == s
+                                 then e `entrySetText` "" >> return False
+                                 else return False
+
+restoreField :: String -> Entry -> EventM EFocus Bool
+restoreField s e = liftIO $ do text <- entryGetText e
+                               if text == ""
+                                  then e `entrySetText` s >> return False
+                                  else return False
+
 
 initTreeView :: TreeView -> TreeStore (EType Error) -> IO ()
 initTreeView view model =  do
@@ -136,43 +157,34 @@ initListView tokens view model = do
     view `treeViewSetModel` model
     view `treeViewSetHeadersVisible` True
 
-    columnTokens <- treeViewColumnNew
-    rendererTokens <- cellRendererTextNew
[…]

### 2009-04-06 src/Annotator/Interface/Handlers.hs

@@ -12,6 +12,7 @@ import Graphics.UI.Gtk
 import Graphics.UI.Gtk.Glade
 import qualified Graphics.UI.Gtk.Gdk.Events as Old
 import System.Glib.Attributes
+import Text.XML.HaXml.XmlContent (XmlContent)
 
 deleteHandler :: Gui -> TreeView -> IO ()
 deleteHandler gui view = do
@@ -44,7 +45,18 @@ addToErrorView gui record@(Record _ (Errtoks tokens) _ _ _) = do
 makeRecord :: Gui -> Error -> IO Record
 makeRecord gui etype = do triggers <- readIORef (trigger gui)
                           tokens  <- readIORef (selectedTkn gui)
-                          return $ Record (Record_Attrs (Just $ token2string triggers) Nothing) (Errtoks $ token2string tokens) etype Nothing Nothing
+                          target <- entryGetText =<< (xmlGetWidget (xml gui) castToEntry "targetField")
+                          comment <- entryGetText =<< (xmlGetWidget (xml gui) castToEntry "commentField")
+                          return $ Record
+                                    (Record_Attrs (Just $ token2string triggers) Nothing)
+                                    (Errtoks $ token2string tokens)
+                                    etype
+                                    (target `truncateIfEqual` "Target Hypothesis" $ Target)
+                                    (comment `truncateIfEqual` "Comment" $ Comment)
+
+truncateIfEqual :: XmlContent a => String -> String -> (String -> a) -> Maybe a
+truncateIfEqual s s' f | s == s'   = Nothing
+                       | otherwise = Just $ f s
 
 highlightTags gui tokens = do
         buffer <- textViewGetBuffer (corpusView gui)

### 2009-04-06 src/Annotator/Interface/Handlers.hs
Wall police.

@@ -58,6 +58,7 @@ truncateIfEqual :: XmlContent a => String -> String -> (String -> a) -> Maybe a
 truncateIfEqual s s' f | s == s'   = Nothing
                        | otherwise = Just $ f s
 
+highlightTags :: Gui -> [String] -> IO ()
 highlightTags gui tokens = do
         buffer <- textViewGetBuffer (corpusView gui)
         table <- textBufferGetTagTable buffer

### 2009-04-06 src/Annotator/Interface.hs
Added sorting by occurrance to cell's tokens.

@@ -24,6 +24,7 @@ import Annotator.Interface.Types
 import Control.Monad (forM,forM_)
 import Control.Monad.Trans (liftIO)
 import Data.IORef
+import Data.List (sort)
 import GHC.List hiding (span)
 import Graphics.UI.Gtk
 import Graphics.UI.Gtk.Gdk.EventM
@@ -175,11 +176,13 @@ initColumn model view f name = do
         view `treeViewAppendColumn` column
 
 record2tokens tokens (Record _ (Errtoks errtoks) _ _ _) =
-        show $ map (tokenString . (tokens!) . read . drop 1) (words errtoks)
+        let sorted = sort $ map (read . drop 1) (words errtoks)
+        in  show $ map (tokenString . (tokens!)) sorted
 record2index (Record _ (Errtoks errtoks) _ _ _) = drop 1 $ head $ words errtoks
 record2trigger tokens (Record (Record_Attrs Nothing _) _ _ _ _) = "N/A"
 record2trigger tokens (Record (Record_Attrs (Just context) _) _ _ _ _) = 
-        show $ map (tokenString . (tokens!) . read . drop 1) (words context)
+        let sorted = sort $ map (read . drop 1) (words context)
+        in  show $ map (tokenString . (tokens!)) sorted
 record2type (Record _ _ etype _ _) = show $ shortenErrorType etype
 record2target (Record _ _ _ Nothing _) = ""
 record2target (Record _ _ _ (Just (Target s)) _) = s

### 2009-04-06 src/Annotator/DTD.hs
Revision of Annotation Scheme.

@@ -30,19 +30,20 @@ newtype Error = Error (Maybe Error_) 		deriving (Eq,Show)
 data Error_ = Error_Context Context
             | Error_Grammar Grammar
             | Error_Spelling Spelling
-            | Error_Gibberish Gibberish
             deriving (Eq,Show)
 newtype Context = Context (Maybe Context_) 		deriving (Eq,Show)
-data Context_ = Context_Lexical Lexical
+data Context_ = Context_Semantics Semantics
               | Context_Idiom Idiom
               | Context_Verbtense Verbtense
+              | Context_Redundancy Redundancy
               deriving (Eq,Show)
 newtype Grammar = Grammar (Maybe Grammar_) 		deriving (Eq,Show)
 data Grammar_ = Grammar_Agreement Agreement
               | Grammar_Voice Voice
-              | Grammar_Valency Valency
-              | Grammar_Complementation Complementation
-              | Grammar_Omission Omission
+              | Grammar_Collocation Collocation
+              | Grammar_Countability Countability
+              | Grammar_Wrong Wrong
+              | Grammar_Omitted Omitted
               | Grammar_Wo Wo
               | Grammar_Morphology Morphology
               | Grammar_Redundancy Redundancy
@@ -62,13 +63,27 @@ data Agreement_ = Agreement_Number Number
                 | Agreement_Person Person
                 | Agreement_Tense Tense
                 deriving (Eq,Show)
-newtype Omission = Omission (Maybe Omission_) 		deriving (Eq,Show)
-data Omission_ = Omission_Subject Subject
-               | Omission_Predicate Predicate
-               | Omission_Object Object
-               | Omission_Article Article
-               | Omission_Preposition Preposition
-               deriving (Eq,Show)
+newtype Omitted = Omitted (Maybe Omitted_) 		deriving (Eq,Show)
+data Omitted_ = Omitted_Subject Subject
+              | Omitted_Predicate Predicate
+              | Omitted_Object Object
+              | Omitted_Noun Noun
+              | Omitted_Conjunction Conjunction
+              | Omitted_Preposition Preposition
+              | Omitted_Pronoun Pronoun
+              | Omitted_Determiner Determiner
+              | Omitted_Adjective Adjective
+              | Omitted_Verb Verb
+              deriving (Eq,Show)
[…]

### 2009-04-06 src/Annotator/Interface/Models.hs

@@ -22,44 +22,59 @@ errorStore = treeStoreNew
         [ leafNode $ EType "Homonymy" (Error  (Just $ Error_Spelling (Spelling (Just $ Spelling_Homonymy Homonymy))))
         , leafNode $ EType "Register" (Error  (Just $ Error_Spelling (Spelling (Just $ Spelling_Register Register))))
         , leafNode $ EType "Diacritics" (Error  (Just $ Error_Spelling (Spelling (Just $ Spelling_Diacritics Diacritics))))
-        , leafNode $ EType "Local" (Error  (Just $ Error_Spelling (Spelling (Just $ Spelling_Local Local))))
+        , leafNode $ EType "Local Variation" (Error  (Just $ Error_Spelling (Spelling (Just $ Spelling_Local Local))))
         ]
-      , Tree.Node (EType "Grammar" (Error  (Just $ Error_Grammar (Grammar  Nothing))))
-        [ Tree.Node (EType "Agreement" (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Agreement (Agreement Nothing))))))
-          [ leafNode $ EType "Number" (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Agreement (Agreement (Just $ Agreement_Number Number))))))
-          , leafNode $ EType "Gender" (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Agreement (Agreement (Just $ Agreement_Gender Gender))))))
-          , leafNode $ EType "Case" (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Agreement (Agreement (Just $ Agreement_Case Case))))))
-          , leafNode $ EType "Person" (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Agreement (Agreement (Just $ Agreement_Person Person))))))
-          , leafNode $ EType "Tense" (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Agreement (Agreement (Just $ Agreement_Tense Tense))))))
+      , Tree.Node (EType "Grammar"           (Error  (Just $ Error_Grammar (Grammar  Nothing))))
+        [ Tree.Node (EType "Agreement"       (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Agreement (Agreement Nothing))))))
[…]

### 2009-04-07 src/Annotator/Interface/Handlers.hs
Some bug fixes, and better place for triggerButton
- actually wipe the text fields when recording something
- do away with the labels on everything
- append to the error list instead of prepending

@@ -39,7 +39,7 @@ recordHandler gui view =
 
 addToErrorView :: Gui -> Record -> IO ()
 addToErrorView gui record@(Record _ (Errtoks tokens) _ _ _) = do
-        (\model -> model `listStorePrepend` record) =<< (readIORef (errorModel gui))
+        (\model -> model `listStoreAppend` record) =<< (readIORef (errorModel gui))
         highlightTags gui (words tokens)
 
 makeRecord :: Gui -> Error -> IO Record
@@ -76,6 +76,8 @@ clearBtnHandler gui =  do writeIORef (selectedTkn gui) []
                           writeIORef (trigger     gui) []
                           writeIORef (currentFocs gui) (selectedTkn gui)
                           (triggerBtn gui) `toggleButtonSetActive` False
+                          (flip entrySetText $ "Target Hypothesis") =<< (xmlGetWidget (xml gui) castToEntry "targetField")
+                          (flip entrySetText $ "Comment") =<< (xmlGetWidget (xml gui) castToEntry "commentField")
                           putTokensOnLabels gui
 
 tagEventHandler :: Gui -> Token -> Old.Event -> TextIter -> IO ()

### 2009-04-07 src/Annotator/Interface/Util.hs

@@ -28,8 +28,12 @@ shortenErrorType = (filter isIrrelevant) . words . (filter $ not.isParen) . show
 putTokensOnLabels :: Gui -> IO ()
 putTokensOnLabels gui = do tkns <- readIORef (selectedTkn gui)
                            trigs <- readIORef (trigger gui)
-                           (tokenLabel gui) `labelSetText` (show $ map tokenString (sort tkns))
-                           (triggerBtn gui) `buttonSetLabel` (show $ map tokenString (sort trigs))
+                           if tkns == []
+                              then (tokenLabel gui) `labelSetText` "Tokens"
+                              else (tokenLabel gui) `labelSetText` (show $ map tokenString (sort tkns))
+                           if trigs == []
+                              then (triggerBtn gui) `buttonSetLabel` "Trigger"
+                              else (triggerBtn gui) `buttonSetLabel` (show $ map tokenString (sort trigs))
 
 tokenId :: Token -> String
 tokenId (Token (Token_Attrs idx) _) = idx

### 2009-04-07 src/Annotator/DTD.hs
Enhanced Taxonomy.
This is the final version.

@@ -36,13 +36,16 @@ data Context_ = Context_Semantics Semantics
               | Context_Idiom Idiom
               | Context_Verbtense Verbtense
               | Context_Redundancy Redundancy
+              | Context_Omitted Omitted
+              | Context_Replace Replace
               deriving (Eq,Show)
 newtype Grammar = Grammar (Maybe Grammar_) 		deriving (Eq,Show)
 data Grammar_ = Grammar_Agreement Agreement
+              | Grammar_Verbtense Verbtense
               | Grammar_Voice Voice
               | Grammar_Collocation Collocation
               | Grammar_Countability Countability
-              | Grammar_Wrong Wrong
+              | Grammar_Replace Replace
               | Grammar_Omitted Omitted
               | Grammar_Wo Wo
               | Grammar_Morphology Morphology
@@ -54,7 +57,6 @@ data Spelling_ = Spelling_Homonymy Homonymy
                | Spelling_Diacritics Diacritics
                | Spelling_Local Local
                deriving (Eq,Show)
-data Gibberish = Gibberish 		deriving (Eq,Show)
 data Lexical = Lexical 		deriving (Eq,Show)
 newtype Agreement = Agreement (Maybe Agreement_) 		deriving (Eq,Show)
 data Agreement_ = Agreement_Number Number
@@ -69,21 +71,23 @@ data Omitted_ = Omitted_Subject Subject
               | Omitted_Object Object
               | Omitted_Noun Noun
               | Omitted_Conjunction Conjunction
+              | Omitted_Adverb Adverb
               | Omitted_Preposition Preposition
               | Omitted_Pronoun Pronoun
               | Omitted_Determiner Determiner
               | Omitted_Adjective Adjective
               | Omitted_Verb Verb
               deriving (Eq,Show)
-newtype Wrong = Wrong (Maybe Wrong_) 		deriving (Eq,Show)
-data Wrong_ = Wrong_Noun Noun
-            | Wrong_Conjunction Conjunction
-            | Wrong_Preposition Preposition
-            | Wrong_Pronoun Pronoun
-            | Wrong_Determiner Determiner
-            | Wrong_Adjective Adjective
-            | Wrong_Verb Verb
-            deriving (Eq,Show)
+newtype Replace = Replace (Maybe Replace_) 		deriving (Eq,Show)
+data Replace_ = Replace_Noun Noun
+              | Replace_Conjunction Conjunction
[…]

### 2009-04-07 src/Annotator/Interface/Models.hs

@@ -36,6 +36,7 @@ errorStore = treeStoreNew
         , leafNode $ EType "Countability"    (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Countability Countability))))
         , leafNode $ EType "Collocation"     (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Collocation Collocation))))
         , leafNode $ EType "Voice"           (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Voice Voice))))
+        , leafNode $ EType "Verb Tense"      (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Verbtense Verbtense))))
         , Tree.Node (EType "Word Order"      (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Wo (Wo Nothing))))))
           [ leafNode $ EType "Internal"      (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Wo (Wo (Just $ Wo_Internal Internal))))))
           , leafNode $ EType "External"      (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Wo (Wo (Just $ Wo_External External))))))
@@ -50,6 +51,7 @@ errorStore = treeStoreNew
           , leafNode $ EType "Predicate"     (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Omitted (Omitted (Just $ Omitted_Predicate Predicate))))))
           , leafNode $ EType "Determiner"    (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Omitted (Omitted (Just $ Omitted_Determiner Determiner))))))
           , leafNode $ EType "Pronoun"       (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Omitted (Omitted (Just $ Omitted_Pronoun Pronoun))))))
+          , leafNode $ EType "Adverb"        (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Omitted (Omitted (Just $ Omitted_Adverb Adverb))))))
           , leafNode $ EType "Preposition"   (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Omitted (Omitted (Just $ Omitted_Preposition Preposition))))))
[…]

### 2009-04-07 src/tools/Agreement.hs
Fixed build.

@@ -11,8 +11,8 @@ import Text.XML.HaXml.XmlContent.Haskell (readXml)
 
 data ComparedRecords = ComparedRecords { agreement    :: [RecordPair]
                                        , disagreement :: ([Record],[Record]) } 
-                                       deriving Show, Eq
-data RecordPair = RecordPair Record Record deriving Eq
+                                       deriving (Show, Eq)
+data RecordPair = RecordPair Record Record deriving (Show, Eq)
 
 main :: IO ()
 main = do args <- getArgs

### 2009-04-07 src/Annotator/Interface/Models.hs
Bumped version, added Replace with X

@@ -63,14 +63,14 @@ errorStore = treeStoreNew
             ]
           ]
         , Tree.Node (EType "Replace"         (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace Nothing))))))
-          [ leafNode $ EType "Determiner"    (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace (Just $ Replace_Determiner Determiner))))))
-          , leafNode $ EType "Pronoun"       (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace (Just $ Replace_Pronoun Pronoun))))))
-          , leafNode $ EType "Adverb"        (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace (Just $ Replace_Adverb Adverb))))))
-          , leafNode $ EType "Preposition"   (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace (Just $ Replace_Preposition Preposition))))))
-          , leafNode $ EType "Noun"          (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace (Just $ Replace_Noun Noun))))))
-          , leafNode $ EType "Conjunction"   (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace (Just $ Replace_Conjunction Conjunction))))))
-          , leafNode $ EType "Adjective"     (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace (Just $ Replace_Adjective Adjective))))))
-          , leafNode $ EType "Verb"          (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace (Just $ Replace_Verb Verb))))))
+          [ leafNode $ EType "with Determiner"with     (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace (Just $ Replace_Determiner Determiner))))))
+          , leafNode $ EType "with Pronoun"with        (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace (Just $ Replace_Pronoun Pronoun))))))
+          , leafNode $ EType "with Adverb"with         (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace (Just $ Replace_Adverb Adverb))))))
[…]

### 2009-04-07 src/Annotator/Interface/Models.hs
Fixed build

@@ -63,14 +63,14 @@ errorStore = treeStoreNew
             ]
           ]
         , Tree.Node (EType "Replace"         (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace Nothing))))))
-          [ leafNode $ EType "with Determiner"with     (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace (Just $ Replace_Determiner Determiner))))))
-          , leafNode $ EType "with Pronoun"with        (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace (Just $ Replace_Pronoun Pronoun))))))
-          , leafNode $ EType "with Adverb"with         (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace (Just $ Replace_Adverb Adverb))))))
-          , leafNode $ EType "with Preposition"with    (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace (Just $ Replace_Preposition Preposition))))))
-          , leafNode $ EType "with Noun"with           (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace (Just $ Replace_Noun Noun))))))
-          , leafNode $ EType "with Conjunction"with    (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace (Just $ Replace_Conjunction Conjunction))))))
-          , leafNode $ EType "with Adjective"with      (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace (Just $ Replace_Adjective Adjective))))))
-          , leafNode $ EType "with Verb"with           (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace (Just $ Replace_Verb Verb))))))
+          [ leafNode $ EType " Determiner"     (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace (Just $ Replace_Determiner Determiner))))))
+          , leafNode $ EType " Pronoun"        (Error  (Just $ Error_Grammar (Grammar  (Just $ Grammar_Replace (Replace (Just $ Replace_Pronoun Pronoun))))))
[…]

### 2009-04-08 src/tools/Agreement.hs
Basic stub of agreement.

@@ -17,38 +17,75 @@ data RecordPair = RecordPair Record Record deriving (Show, Eq)
 main :: IO ()
 main = do args <- getArgs
           case args of
-
-              files@(_:_:_) -> do parses <- forM files (\f -> fmap readXml (readFile f))
-                                  case lefts parses of
-                                       [] -> compareCorpora $ rights parses
-                                       errors -> do forM_ errors putStrLn
+              (_:_:_) -> treatFiles args
               _       -> putStrLn "Usage: agreement a1.xml a2.xml"
+
+treatFiles :: [FilePath] -> IO ()
+treatFiles files = do parses <- forM files (\f -> fmap readXml (readFile f))
+                      case lefts parses of
+                           [] -> compareCorpora $ rights parses
+                           errors -> do forM_ errors putStrLn
+
 
 compareCorpora :: [Corpus] -> IO ()
 compareCorpora ((Corpus (Tokens a1 ts1) (Errors es1)):(Corpus (Tokens a2 ts2) (Errors es2)):[]) = do
     putStrLn "Testing for token to token Consistency..."
-    if a1 /= a2 || ts1 /= ts2
+    if a1 /= a2 -- || ts1 /= ts2
        then putStrLn "Tokens inconsistent. Checking agreement doesn't make sense this way."
        else do putStrLn "Tokens consistent."
-               let results = compareErrors2 es1 es2
-               undefined
+               compareErrors2 es1 es2
 compareCorpora _          = putStrLn "Sorry, but only two corpora can be evaluated at this point."
 
-compareErrors2 :: [Record] -> [Record] -> ComparedRecords
-compareErrors2 es1 es2 = let ses1 = sortBy tokenEquality es1
-                             ses2 = sortBy tokenEquality es2
-                         in trawlErrors ses1 ses2 (ComparedRecords [] ([],[]))
+compareErrors2 :: [Record] -> [Record] -> IO ()
+compareErrors2 es1 es2 = do
+        printColumn2 "Annotations" (show $ length es1) (show $ length es2)
+        printColumn "Total agreement" (show . length $ intersectBy recordIdentity es1 es2)
+        printColumn "Token agreement" (show . length $ intersectBy tokenIdentity es1 es2)
[…]

### 2009-04-08 text/paper.tex
some work on paper

@@
 \usepackage[utf8x]{inputenc}
+\usepackage{aleks}
 \author{Magdalena Leshtanska \and Aleksandar Dimitrov}
@@
 \maketitle
+\tableofcontents
+\abstract{I am an abstract!}
 \section{Introduction}\label{sec:intro}
@@
 Based on these assumptions, we decided to decouple the \textit{error markup}
-from the \textit{corpus data}. Specifically, to our annotation scheme, the
+from the \textit{corpus data}. Specifically, to our annotation scheme the
 errors and the text are two entirely different data structures. Every
@@
 similar to the way modern Relational Data Bases reference their data. This makes
-the index more accessible and easier to maintain. Every marked up error contains exactly
-the amount of information it needs in order to justify its presence, viz. by
-defining what triggered its occurrence, and sufficient classification within the
-taxonomy presented in \ref{sec:taxo}.
+the index more accessible and easier to maintain.
 
@@
 meticulously well-defined manual is still going to be read and applied by
-humans, who all have their personal intuition and character. It is easy to see
+humans, who will have their personal intuition. It is easy to see
 why this inherent individuality poses problems to the consistency of any kind of
-annotated data. A well-written annotation manual can limit the effect this
-natural variance can have on the data, but applying measures of consistency is
-strictly necessary. The assessment of our data will be discussed in section
-\ref{sec:results} while the remainder of this section will explain the rationale
-behind our annotation manual, which is in turn given in section \ref{sec:manual}.
+annotated data. A well-written annotation manual will limit the effect this
+natural variance can have on the data.
 
-\subsection{Calling Errors by Name}\label{sec:taxo}
+\subsection{Calling Errors by Name}\label{sec:taxonomy}
 
@@
 
-Furthermore, our taxonomy will have a base case, called \textbf{gibberish} for
-every error that cannot otherwise be classified, necessitated by its
-deliberately established incompleteness. Sections delineated as such a case
-should, however, be few in number, and as local as possible.
-
-There are three other kinds of errors we want to term:
+Our annotation scheme presents three distinct kinds of errors.
 
@@
 should have ended up as.
-\item \textit{Form errors} are otherwise grammatical sentences which are,
+\item \textit{Context errors} are otherwise grammatical sentences which are,
 however, semantically incorrect given the current context.
@@
 
-\subsection{Representing Errors in Annotations}
+\subsection{Representing Errors in Annotations}\label{sec:errorformat}
+
+An error annotation is a tuple $\pair{E,C,\theta,t,c}$, where
+\begin{itemize}
+  \item $E$ is a nonempty set of indices of erroneous tokens,
+  \item $C$ is a possibly empty set of context tokens 
+  \item $\theta$ is the type of the error,
+  \item $s$ a string denoting an optional target hypothesis hint, and
+  \item $c$ an optional comment.
+\end{itemize}
+
+The error type $\theta$ is defined as an ordered sequence of categories from the
+taxonomy presented in \ref{sec:taxonomy}.
+
+We hope this general format to be sufficient for annotation within our current
+goals, and extensible enough to cover future refinements of the annotation
+standard, such as the addition of part of speech tags or syntactic markup.
+Keeping the error annotation in a separate data structure, allows for the
+annotation itself to be more detailed and flexible.
 
@@
 
-\subsection{Form Errors}
-Form errors occur in otherwise completely grammatical sentences, which, however
+\subsection{Context Errors}
+Context errors occur in otherwise completely grammatical sentences, which, however
 do not convey what was obviously intended by the author taking context into
-account. Lexical errors are also form errors, if the given word is grammatically
+account. Lexical errors are also context errors, if the given word is grammatically
 valid within the sentence.
@@
 
-\section{Results, Assessment and Future Work}\label{sec:results}
+\section{Conclusion}\label{sec:results}
+
+Statistical inter-annotator agreement  measures are a common quality assessment
+method used in corpus linguistics and related fields. Hereby, annotations made
+on a particular data set by two or more annotators are compared using
+quantitative methods.  \cite{ap2008} give an overview of currently employed
+methods.
+
+While inter-annotator agreement measures have been applied successfully to
+various corpus linguistic tasks, so far they have not found wide usage among
+learner language annotation. In fact, we believe the current techniques  are not
+applicable to this particular problem domain.
+
+\subsection{Unitization and Multidimensional Markup}
+
+Existing inter-annotator agreement measures %such as … \cite FIXME all assume
+the presence of atomic units in the corpus data, which are annotated by all
+annotators of a certain data set. This is the case with part of speech tagging
+or syntactic classification of chunked sentences, where all annotators are
+presented with a set of tokens or a set of chunks, respectively, which they have
+to annotate.
+
+The annotations \textit{over these units} are then used to calculate an
+agreement coefficient that will usually range from -1 to 1. However, learner
+language is a lot more diverse in nature, and it is not as easily possible to
+partition it into atomic data units. Instead, annotators have to \textit{find
+data points} and also \textit{manually delineate} them, according to the
+annotation manual. This brings the problem of \textit{unitization} into play.
+
+\cite{ap2008} briefly discuss unitization and go on to note that it has thus far
+not been exhaustively researched. Even more importantly, they explicitly comment
+on the unknown status of the validity of the only inter-annotator agreement
+measure in the corpus linguistic literature, $\alpha_U$, presented in
+\cite{krip1995}.
+
+Apart from being untested, $\alpha_U$ has several more problems that make it an
+ill fit for learner language data. As mentioned in \cite{ap2008}, Krippendorf's
+$\alpha_U$ assumes markup to be non-overlapping. This is certainly not the case
+with learner language, where one error may be nested within in another
+error\footnote{This happens very frequently with spelling errors.}. Also, if a
+segment annotated by one annotator spans more than one segment annotated by the
+other annotator, $\alpha_U$ will not calculate their agreement correctly.
+
+We see yet another problem with applying $\alpha_U$ to learner corpora: we
+showed earlier % FIXME: where
+that often learner errors will not present a coherent unit, but may be scattered
+across several tokens. $\alpha_U$ does not account for the possibility of
+annotations deviating \textit{within} the boundaries of a marked up error.
+
+\subsection{Possible Extensions}
+
+After assessing the quality of our data, we reached
+the conclusion that the error format described in \ref{sec:errorformat}
+might benefit from several refinements. Adding a field for part of speech tags
+might contribute to the clarity of the data, as well as to its searchability.
+The annotated corpus could be queried for erroneously placed verbs or
+prepositions, for example.
+
+Moreover, we came to the conclusion that defining the error context as a set of
+tokens might be misleading or at least difficult to understand in case the error
+context does not constitute one sequence, but several scattered sequences, such
+as proper nouns or syntactic constituents. $C$ could therefore be turned into a
+set of sequences of token indices.
+
+\subsubsection{Towards an underspecification formalism for target hypotheses}
+
+During the annotation process, we discovered that our taxonomy branches for
+\textit{omission}, \textit{replacement}, and \textit{redundancy} could be turned
+into a stub of a formalism for underspecification of target hypotheses. Instead
+of giving a string for a target hypothesis, such a formalism would make it
+possible to approximate the target and therefore allow for more flexibility in
+the markup. Note, however, that these forms of omission, redundancy and
+replacement differ from the ones included in the error taxonomy.
+
+The taxonomy tries to account for \textit{what is wrong} with a given string of
+text. A target hypothesis would try to make assumptions about \textit{how this
+could be fixed}. Our categories in the error taxonomy suggesting manipulation of
+the input text were explicitly designed to catch cases where a clear reason for
+an error could not be found, and the syntactic environment of a given set of
+tokens would require the text to be changed entirely. Such subcategorization
+mistakes could be granted their own category and the target hypothesis could
+account for the necessary steps in order to ensure grammaticality.
+
+This would also allow for existing annotations to be combined with
+\textit{generic instructions for correcting the input} and increase the
+granularity of the data.
 

### 2009-04-08 src/Annotator/Interface/Handlers.hs
Added rudimentary change button.

@@ -14,6 +14,19 @@ import qualified Graphics.UI.Gtk.Gdk.Events as Old
 import System.Glib.Attributes
 import Text.XML.HaXml.XmlContent (XmlContent)
 
+--changeButtonHandler :: Gui -> IO ()
+changeButtonHandler gui etypeview elist = do
+    [path] <- treeSelectionGetSelectedRows =<< (treeViewGetSelection etypeview)
+    (EType _ record) <- (\store -> store `treeStoreGetValue` path) =<< errorStore
+
+    lmodel <- readIORef (errorModel gui)
+    [row] <- treeSelectionGetSelectedRows =<< (treeViewGetSelection elist)
+    (Record attrs tokens _ target comment) <- lmodel `listStoreGetValue` (head row)
+    deleteThisRecord gui row
+    lmodel `listStoreAppend` (Record attrs tokens record target comment)
+    return ()
+
+
 deleteHandler :: Gui -> TreeView -> IO ()
 deleteHandler gui view = do
     rows <- treeSelectionGetSelectedRows =<< (treeViewGetSelection view)

### 2009-04-08 src/Annotator/Interface/Handlers.hs
Fixed potential crash.

@@ -16,15 +16,20 @@ import Text.XML.HaXml.XmlContent (XmlContent)
 
 --changeButtonHandler :: Gui -> IO ()
 changeButtonHandler gui etypeview elist = do
-    [path] <- treeSelectionGetSelectedRows =<< (treeViewGetSelection etypeview)
-    (EType _ record) <- (\store -> store `treeStoreGetValue` path) =<< errorStore
-
+    paths <- treeSelectionGetSelectedRows =<< (treeViewGetSelection etypeview)
     lmodel <- readIORef (errorModel gui)
-    [row] <- treeSelectionGetSelectedRows =<< (treeViewGetSelection elist)
-    (Record attrs tokens _ target comment) <- lmodel `listStoreGetValue` (head row)
-    deleteThisRecord gui row
-    lmodel `listStoreAppend` (Record attrs tokens record target comment)
-    return ()
+    rows <- treeSelectionGetSelectedRows =<< (treeViewGetSelection elist)
+
+    case paths of
+         [path] -> case rows of
+                        [row] -> do (EType _ record) <- (\store -> store `treeStoreGetValue` path) =<< errorStore
+                                    (Record attrs tokens _ target comment) <- lmodel `listStoreGetValue` (head row)
+                                    deleteThisRecord gui row
+                                    lmodel `listStoreAppend` (Record attrs tokens record target comment)
+                                    return ()
+                        _ -> putStrLn "No row seleceted"
+         _ -> putStrLn "No path seleceted"
+
 
 
 deleteHandler :: Gui -> TreeView -> IO ()

### 2009-04-08 src/tools/Agreement.hs
Some further changes.

@@ -39,32 +39,62 @@ compareCorpora _          = putStrLn "Sorry, but only two corpora can be evaluat
 compareErrors2 :: [Record] -> [Record] -> IO ()
 compareErrors2 es1 es2 = do
         printColumn2 "Annotations" (show $ length es1) (show $ length es2)
-        printColumn "Total agreement" (show . length $ intersectBy recordIdentity es1 es2)
-        printColumn "Token agreement" (show . length $ intersectBy tokenIdentity es1 es2)
-        printColumn "token + subtree" (show . length $ intersectBy tokenSubtree es1 es2)
-        printColumn "token subset" (show . length $ intersectBy tokenSubset es1 es2)
-        printColumn "subset subtree" (show . length $ intersectBy subsetSubtree es1 es2)
-        printColumn "nonmatches" (show . length $ intersectBy (not subsetSubtree) es1 es2)
+        putStrLn "\nThree data points\n***"
+        printColumn "io + iy + ir" $
+                evaluate ((equality errtoks) /&\ (equality etype) /&\ (equality trigger)) es1 es2
+        printColumn "so + iy + ir" $
+                evaluate ((subset errtoks) /&\ (equality etype) /&\ (equality trigger)) es1 es2
+        printColumn "io + sy + ir" $
+                evaluate ((equality errtoks) /&\ (subtree etype) /&\ (equality trigger)) es1 es2
+        printColumn "io + iy + sr" $
+                evaluate ((equality errtoks) /&\ (equality etype) /&\ (subset trigger)) es1 es2
+        printColumn "so + sy + ir" $
+                evaluate ((subset errtoks) /&\ (subtree etype) /&\ (equality trigger)) es1 es2
+        printColumn "io + sy + sr" $
+                evaluate ((equality errtoks) /&\ (subtree etype) /&\ (subset trigger)) es1 es2
+        printColumn "so + iy + sr" $
+                evaluate ((subset errtoks) /&\ (equality etype) /&\ (subset trigger)) es1 es2
+        printColumn "so + sy + sr" $
+                evaluate ((subset errtoks) /&\ (subtree etype) /&\ (subset trigger)) es1 es2
+        putStrLn "Two data points\n***"
+        printColumn "io + iy" $
[…]

### 2009-04-08 text/paper.tex

@@
-\documentclass[12pt]{article}
+\documentclass{scrartcl}
 \usepackage[utf8x]{inputenc}
@@
 
+[Follows information about triggers, and how the play into this general scheme ]
+
 \subsection{Representing Errors in Annotations}\label{sec:errorformat}
@@
 
-\section{Conclusion}\label{sec:results}
+\section{Assessment}\label{sec:results}
 
@@
 
+\subsection{Quantificational Analysis}
+
+\begin{table}
+  \centering
+\begin{tabular}{r|c|c}
+  Total: & (905,933)& (98.69\%,101.74\%)\\\hline
+  $io \wedge iy \wedge ir$:& 438 & 47.76 \%\\
+  $so \wedge iy \wedge ir$:& 448 & 48.85 \%\\
+  $io \wedge sy \wedge ir$:& 447 & 48.74 \%\\
+  $io \wedge iy \wedge sr$:& 480 & 52.34 \%\\
+  $so \wedge sy \wedge ir$:& 476 & 51.90 \%\\
+  $io \wedge sy \wedge sr$:& 503 & 54.85 \%\\
+  $so \wedge iy \wedge sr$:& 499 & 54.41 \%\\
+  $so \wedge sy \wedge sr$:& 544 & 59.32 \%\\
+  $     io \wedge iy$:& 486      & 52.99 \%\\
+  $     so \wedge iy$:& 506      & 55.17 \%\\
+  $     io \wedge sy$:& 509      & 55.50 \%\\
+  $     so \wedge sy$:& 551      & 60.08 \%\\
+  $     io \wedge ir$:& 525      & 57.25 \%\\
+  $     so \wedge ir$:& 578      & 63.03 \%\\
+  $     io \wedge sr$:& 623      & 67.93 \%\\
+  $     so \wedge sr$:& 719      & 78.40 \%\\
+  $          io$:& 629           & 68.59 \%\\
+  $          so$:& 730           & 79.60 \%\\
+\end{tabular}
+\caption{Absolute amount of annotation overlap.}\label{tab:1}
+
+\end{table}
+
+Table \ref{tab:1} shows the total amount of corresponding data in the individual
+markup. The table's labeling reads as follows: $s$ stands for a non-empty
+intersection (or prefix relation for the error types) and $i$ for equality.
+$o,y,r$ denote the data type: $o$ stands for error tOkens, $y$ for the error's
+tYpe, and $r$ for the errors context (or \textit{tRigger}). Thus, $so\wedge
+sy\wedge ir$ is the number of all annotations that have a partial overlap on the
+error tokens $E$, a prefix match on the error type $\theta$, and identical error
+context.
+
 \subsection{Possible Extensions}

### 2009-04-09 text/paper.tex
Tons of work on the paper...

--- 2009-04-08 text/paper.tex
@@
-\documentclass[12pt]{article}
+\documentclass[12pt]{scrartcl}
 \usepackage[utf8x]{inputenc}
-\usepackage{aleks}
+\usepackage{graphicx,linguex,natbib,aleks,amsmath,amsthm}
 \author{Magdalena Leshtanska \and Aleksandar Dimitrov}
@@
 \maketitle
+
+\newtheorem{thm}{Theorem}[section]
+\newtheorem{definition}[thm]{Definition}
+
 \tableofcontents
@@
 \item \textit{Interleaving annotations} occur when an error doesn't end before
-another one begins. Given the tokens $a_1 a_2 a_3$, and two errors $e_1$ and
-$e_2$ ranging over tokens $a_1$, $a_2$ and $a_2$ $a_3$ respectively, the
-resulting markup will be confusing or outright impossible to read: $(e_1)a_1
-(e_2)a_2(/e_1)a_3(/e_2)$. This is particularly a problem with XML,
+another one begins. Given the tokens $\tau_1 \tau_2 \tau_3$, and two errors
+$\eta_1$ and
+$\eta_2$ ranging over tokens $\tau_1$, $\tau_2$ and $\tau_2$ $\tau_3$ respectively, the
+resulting markup will be confusing or outright impossible to read: $(\eta_1)\tau_1
+(\eta_2)\tau_2(/\eta_1)\tau_3(/\eta_2)$. This is particularly a problem with XML,
 since the specification\footnotemark explicitly disallows interleaving
@@
 \item \textit{Greedy annotation} covers tokens entirely uninvolved in the
-\textit{``production''} of an error. If of the tokens $a_1 a_2 a_3$ only $a_1$
-and $a_3$ are involved in the production of an error (say, an agreement error)
-marking the entire token sequence as erroneous would involve blaming the
-otherwise completely innocent token $a_2$. This can significantly increase the
-noise in a corpus.
+\textit{``production''} of an error. If of the token sequence $\tau_1 \tau_2
+\tau_3$ only $\tau_1$
+and $\tau_3$ are erroneous,
+marking the entire token sequence with an error annotation would falsely accuse
+the
+otherwise completely innocent token $\tau_2$.
 \end{itemize}
@@
 
-There is little data about the primary use cases for learner corpora known to
-us, but it does not seem too far fetched to assume most linguists interested in
-such data are not likely to browse through the corpora in order to find an
-amusingly written piece of text. Instead, it seems to be safe to assume that
-most would be interested in the \textbf{kinds of errors} that can occur in such
-a corpus, and their particular properties. The error data should thus contain as
-little noise as possible and be maximally specific while retaining a large amount
-of generality.  It should furthermore be easily accessible, without having to
-actually read the texts within the corpus itself.
-
 Based on these assumptions, we decided to decouple the \textit{error markup}
-from the \textit{corpus data}. Specifically, to our annotation scheme the
+from the \textit{corpus data}. Specifically, to our annotation method, the
 errors and the text are two entirely different data structures. Every
@@
 similar to the way modern Relational Data Bases reference their data. This makes
-the index more accessible and easier to maintain.
+the index more accessible and easier to maintain and eliminates both problems
+above, because every single annotation can be completely independent from all
+other annotations.
 
 \section{Error Taxonomy and Annotation Scheme}\label{sec:scheme}
-
-Learner errors are not easily classified due to their wide variety of
-appearance, kind, and ambiguities involved. Typically, an annotation scheme will
-strive to cover as large a scope it possibly can unambiguously classify, yet
-there are many practical limitations to the potential coverage one can achieve.
-
-One of the most prevalent restrictions is the need to devise the annotation
-manual with great attention to detail, and to perform a meaningful assessment of
-the quality of the annotated data after the annotation process. Even a most
-meticulously well-defined manual is still going to be read and applied by
-humans, who will have their personal intuition. It is easy to see
-why this inherent individuality poses problems to the consistency of any kind of
-annotated data. A well-written annotation manual will limit the effect this
-natural variance can have on the data.
-
-\subsection{Calling Errors by Name}\label{sec:taxonomy}
 
@@
 
-Our annotation scheme presents three distinct kinds of errors.
+\subsection{Basic Concepts}\label{sec:threeholies}
 
-\begin{enumerate}
-\item \textit{Spelling errors} are a primitive form of error. They contain no
-meta-data or any kind of further specification except maybe a target word they
-should have ended up as.
-\item \textit{Context errors} are otherwise grammatical sentences which are,
-however, semantically incorrect given the current context.
-\item \textit{Grammar errors} simply do not belong to the target language because
-of their erroneous morphology or syntax.
-\end{enumerate}
+Our annotation scheme presents three distinct kinds of errors:
 
-\subsection{Representing Errors in Annotations}\label{sec:errorformat}
+\begin{itemize}
+  \item \textit{Grammatical context.} each introduced token enforces constraints
+  on the utterance, or on parts of it. In \ref{swim}, \textit{remember} requires
+  a gerund, thus the form \textit{swim} is ungrammatical in this context. We
+  call such mistakes \textbf{grammar errors}.
 
-An error annotation is a tuple $\pair{E,C,\theta,t,c}$, where
+  \ex. *I remember swim in the river\label{swim}
+
+  \item \textit{Semantic context.}  In \ref{glue}, though the grammatical
+  structure is correct, the predicate does not fit the contextual information.
+  This is a contextual error.
+
+  \ex. *Yesterday I will go to the fair.\label{glue}
+
+  \item \textit{Spelling errors.} Spelling errors are typically not influenced
+  by contextual information, except for the normative context of a given
+  language's orthographic rules. Punctuation is a special case, since it may
+  also carry semantic information.
+
+  \end{itemize}
+
+\subsection{Calling Errors by Name}\label{sec:taxonomy}
+
+The classification of errors, as well as establishing an error's scope and 
+influence on the rest of the sentence are the main tasks in the annotation
+process of learner language. The following section documents our approach to
+these problems.
+
+\subsection{Error Context}
+
+Annotated learner corpora serve one primary purpose: categorizing and
+cataloguing different kinds of errors that may occur in learner language. In
+order for such data to be maximally useful, the error annotations have to be as
+general as possible. We strive to improve the quality of the data by only
+annotating erroneous tokens in a given error tag, thus not erroneously catching
+'good' tokens in our error annotation. This, however, comes at a cost: if only
+the ungrammatical token is identified as an error, its classification is no
+longer justified from with in the error annotation itself. Consider the
+following sentence:
+
+\ex. Before she came, *She had going to the super market.
+
+Here, the only erroneous token is \textit{going}, yet it is not by itself a
+wrong word. This is where error context comes into play: since the error
+annotation could not possibly tag "going" as a "badly formed verb tense", because, by
+itself there is nothing wrong with it, we add "Before she came" to its
+\textbf{error context} and apply the aforementioned type to the whole unit
+consisting of erroneous token and error context. This error dependency mechanism
+allows error annotations to be confined to a minimal space and still be
+interpretable by themselves.
+
+Another common pitfall for error taxonomies are ambiguous cases where the
+annotator has to decide between several possible annotations. This may be
+avoided by advising annotation of every conceived alternative in such cases,
+creating overlapping markup. Typically, morphological errors, spelling errors,
+word order errors, and other kinds of errors will form distinct levels that can
+stack to a cumulative layer of errors on a single token. It is also interesting
+to note that context/error pairs can form locking formations, where one error's
+context can be another error, whose context will point to the original error in
+turn.
+
+\ex. *A mobile phones can be very useful.
+
+Here, the determiner and the subject do not agree in number, but it is not clear
+which one is wrong. There are two possible corrections, which would result
+in the subject ending up as plural or singular, respectively. Therefore, we tag
+two errors, a number agreement error on \textit{a} with context \textit{phones}, and
+one on \textit{phones}, with context \textit{a}. Note that annotating only "a
+phones" as an agreement error would not account for the two possible target
+hypotheses. Such a scheme would have to invent a mechanism for defining multiple
+target hypotheses\footnote{Although it is not clear whether our approach would
+rid us of the necessity of such mechanisms. Word order mistakes pose a
+significant problem to a one-target-hypothesis-per-error approach.}.
+
+%In fact, in this way, the process of annotation could be partly automated, using an HPSG style feature compatibility check, adjusted for the purpose -  e.g.
+removing head dominance assumtpions (because in learner language, the head itself may be wrong), and instead taking a linear processing approach, to determine where grammatical errors are.
+
+\subsection{Error Taxonomy}
+We adopted an error taxonomy to enrich our error identification annotations by
+additional error classification. Error type is not dependent on the trigger
+policy, any grammatical error type annotation scheme may be used here. To adjust
+an annotation scheme to the error context paradigm, one must simply divide the
+types among our three basic error type categories: Spelling, Grammar and
+Context, as already described in \ref{sec:threeholies}.
+
+For our test taxonomy, we used the taxonomy described in \cite{negrillo} as a
+base, and enhanced it by several concepts from the CLC annotation scheme, as
+specified in \cite{CLC}. While we retained the general ideas and structure of
+the aforementioned schemata, we aimed to improve their generality, and
+systematization.
+
+\begin{figure}
+\centering
+\includegraphics[scale=.2]{taxonomy}
+\caption{Error classification taxonomy.}
+\end{figure}
+
+We chose to follow a hierarchical setup, with some emphasis on typical decision
+paths an annotator will have to make during the annotation process. The
+hierarchical nature of our taxonomy made it a good fit for XML schemas, which
+define relationships in similar ways.
+
+In this taxonomy, "Error" is the root of all types of errors, with more and more
+specific error classifications percolating down the tree. An annotator may
+choose a less specific category in case they are unsure about a certain item.
+Some errors can indeed not be classified at all, and will have to remain
+annotated as the general category "Error." Since annotation of learner language
+can be tricky at times, we chose to allow these kinds of underspecification of
+error class in order to give an annotator the possibility to express an error
+type more flexibly.
+
+We also chose to alter a few concepts of the error taxonomies we based our own
+on, specifically, we chose to eliminate certain kinds of error types to make the
+annotation task less involved and more accurate.
+
 \begin{itemize}
-  \item $E$ is a nonempty set of indices of erroneous tokens,
-  \item $C$ is a possibly empty set of context tokens 
+  \item \textit{Normative Errors.} Some normative concepts are present in almost
+  all European languages, such as capitalization at the beginning of a sentence.
+  We chose to ignore such errors, since they are not a sign of weakness in the
+  target language but likely pure by chance mistakes.
+  \item \textit{Style Errors.} Style is too soft a concept. Errors in style are
+  notoriously hard to peg and reliably quantify. Moreover, the NOCE corpus
+  contained data mostly from beginning English learners, where style errors are
+  not as important, evident or relevant.
+  \item \textit{Punctuation.} Punctuation mistakes are a very delicate concern:
+  they can have tremendous effects on any given token string on both a syntactic
+  and semantic level. They might change constituent boundaries, sentence
+  boundaries, argument structures and many more things. Moreover, they tend to
+  have cascading effects on the correctness of a given sentence. Again, because
+  we were annotating a beginner's corpus, we chose to ignore these kinds of
+  errors, since in beginner language punctuation often degrades to line noise
+  very quickly.
+\end{itemize}
+
+We also included a "transfer" attribute to indicate that an error is an $L_1$
+transfer error. This attribute was, however, not used in the annotation process,
+since the annotators' knowledge of the Spanish language was insufficient to
+make reliable judgements about it.
+
+\subsection{Markup}
+
+The source files to our corpus consisted of plain text files from the
+NOCE-corpus % FIXME cite
+of beginning to intermediate Spanish learners of English. Our corpus
+format is defined in \ref{def:errorformat}
+
+\begin{definition}\label{def:errorformat}
+
+A corpus $\mathcal{C}=\pair{\mathcal{T},\mathcal{E}}$ is defined as a pair of
+a set of tokens, $\mathcal{T}$ and a 
+set of errors, $\mathcal{E}$. Every token $\tau_i$ is indexed with a unique
+identifier subscript $i$\footnote{Note that the identifier may not consist entirely of
+numbers, since it's XML type is \texttt{ID}, which demands identifiers to be
+alphanumeric sequences. Furthermore, the identifiers need not adhere to any
+particular order, as long as they are unique.}.
+An error annotation is a tuple $\eta$ = $\pair{E,C,\theta,t,c}$, where
+\begin{itemize}
+  \item $E \subseteq \mathcal{T}$ is a nonempty set of indices of erroneous tokens
+  \item $C \subseteq \mathcal{T}$ is a possibly empty set of context tokens
   \item $\theta$ is the type of the error,
@@
 The error type $\theta$ is defined as an ordered sequence of categories from the
-taxonomy presented in \ref{sec:taxonomy}.
+taxonomy presented in \ref{sec:taxonomy}. Furthermore, $E \cap C = \emptyset$
+for all $\eta\in\mathcal{E}$.
 
-We hope this general format to be sufficient for annotation within our current
-goals, and extensible enough to cover future refinements of the annotation
-standard, such as the addition of part of speech tags or syntactic markup.
-Keeping the error annotation in a separate data structure, allows for the
-annotation itself to be more detailed and flexible.
+\end{definition}
 
-\section{Annotation Manual}\label{sec:manual}
+This general data structure is translated to XML. A DTD ensures correct usage of
+the markup. The authors also designed a graphical user interface based
+annotation tool\footnotemark which hides the implementation details of the
+corpus data from the user and enables a rapid annotation process. The user marks
+tokens from the corpus as erroneous, assigns them a type, possibly an error
+context and suggest a target hypothesis or records a comment entirely via the
+interface.
 
-\subsection{Spelling Errors}\label{sec:speling}
-A spelling error is an error that \textit{cannot} be explained by mistakes in
-morphology, e.g. derivation or inflection. It will typically comprise only one
-word. If it comprises more than one word, it is seen as a \textit{compound}
-spelling error, i.e. two tokens separated by a space should have been
-concatenated to yield only one token. Similarly, if the target hypothesis
-accounts for more than one word, the token should have been split up into two.
+\footnotetext{The tool is implemented in Haskell (\texttt{www.haskell.org}) on
+top of the GTK  framework (\texttt{www.gtk.org}). It free and open source. The
+sources to the tool and the XML markup format are given in appendix
+\ref{appendix:github}.}
 
-\subsection{Context Errors}
-Context errors occur in otherwise completely grammatical sentences, which, however
-do not convey what was obviously intended by the author taking context into
-account. Lexical errors are also context errors, if the given word is grammatically
-valid within the sentence.
+\subsection{Limitations}
+Some elements, for example word order errors have enforced context, but it is hard to express via our schema, so we don't do it. 
+No correct positions specified by word order errors - also hard to express.
+One could assign them a target position after an existing token, but then one must also keep track of other "reorderings" by multiple word order errors - I would like beer some to drink - here, both erroneous units - some and to drink should be placed between like and beer, but one must take care to order them correctly.
+Also, as our schema tags all possible errors, in the case of interchangeable errors (cf XXX), one cannot be certain if the error annotated is really the one, made by the learner, and not the other possible one.
+Most  importantly, there are sometimes more than one places where a word sequence can fit in.
 
-\subsection{Grammatical Errors}
+Errors within a token cannot be analyzed – for example, agglutination, because a token is our smallest unit.
 
-\subsection{Devising Target Hypotheses}
-Retain existing input by the author of the text to the greatest extent possible.
-The text should be only minimally altered in the target hypothesis.
-
-\section{Conclusion}\label{sec:results}
+\section{Assessment}\label{sec:results}
 
@@
 various corpus linguistic tasks, so far they have not found wide usage among
-learner language annotation. In fact, we believe the current techniques  are not
+learner language annotation. We believe the current techniques  may not
 applicable to this particular problem domain.
@@
 
-Existing inter-annotator agreement measures %such as … \cite FIXME all assume
-the presence of atomic units in the corpus data, which are annotated by all
-annotators of a certain data set. This is the case with part of speech tagging
-or syntactic classification of chunked sentences, where all annotators are
-presented with a set of tokens or a set of chunks, respectively, which they have
-to annotate.
-
-The annotations \textit{over these units} are then used to calculate an
-agreement coefficient that will usually range from -1 to 1. However, learner
-language is a lot more diverse in nature, and it is not as easily possible to
-partition it into atomic data units. Instead, annotators have to \textit{find
-data points} and also \textit{manually delineate} them, according to the
-annotation manual. This brings the problem of \textit{unitization} into play.
+Existing inter-annotator agreement measures all assume the presence of atomic
+units in the corpus data, which are annotated by the annotators of a certain
+data set.  The annotations \textit{over these units} are then used to calculate an
+agreement coefficient. However, constituents of learner language, being fairly
+diverse in nature, are not as easily contained in atomic units. Instead, a
+common task in the annotation of learner language is \textit{delineating}
+the extent of a certain mistake.
 
@@
 measure in the corpus linguistic literature, $\alpha_U$, presented in
-\cite{krip1995}.
+\cite{krip1995}.  Apart from being untested, $\alpha_U$ also seems to have
+problems with overlapping markup both in the text, and between annotation sets
+(both of which are frequent in learner language data.) The measure also assumes
+annotation spans to be continuous, which is not the case in our data.
 
-Apart from being untested, $\alpha_U$ has several more problems that make it an
-ill fit for learner language data. As mentioned in \cite{ap2008}, Krippendorf's
-$\alpha_U$ assumes markup to be non-overlapping. This is certainly not the case
-with learner language, where one error may be nested within in another
-error\footnote{This happens very frequently with spelling errors.}. Also, if a
-segment annotated by one annotator spans more than one segment annotated by the
-other annotator, $\alpha_U$ will not calculate their agreement correctly.
+\subsection{Quantificational Analysis}
 
-We see yet another problem with applying $\alpha_U$ to learner corpora: we
-showed earlier % FIXME: where
-that often learner errors will not present a coherent unit, but may be scattered
-across several tokens. $\alpha_U$ does not account for the possibility of
-annotations deviating \textit{within} the boundaries of a marked up error.
+Facing these theoretical difficulties, we reached the conclusion that the only
+viable way to define inter-annotator agreement over learner language data
+we could improvise\footnotemark would still not yield interpretable results. In
+order to quantify our analysis efforts, we analyzed the two annotation sets with
+respect to their annotation's intersections.
+
+\footnotetext{It would in theory be possible to regard tokens as units, and that
+is indeed how our scheme is currently implemented. An agreement measure would
+then interpret annotations locally on these units only. However, this method has
+two major drawbacks: firstly, it does not account for more than one annotation
+on a given unit, which happens frequently. Also, if one annotator marks a set of
+tokens as an error consisting of more than one token, and the other annotator
+marks these tokens with several errors of the same category (which can happen,
+for example, with word order mistakes, as well as complex agreement mistakes),
+this method would fail to account for the discrepancy.}
+
+\begin{table}
+  \centering
+\begin{tabular}{r|c|c}\ref{tab:1}
+  Total: & (905,933)& (98.69\%,101.74\%)\\\hline
+  $io \wedge iy \wedge ir$:& 438 & 47.76 \%\\
+  $so \wedge iy \wedge ir$:& 448 & 48.85 \%\\
+  $io \wedge sy \wedge ir$:& 447 & 48.74 \%\\
+  $io \wedge iy \wedge sr$:& 480 & 52.34 \%\\
+  $so \wedge sy \wedge ir$:& 476 & 51.90 \%\\
+  $io \wedge sy \wedge sr$:& 503 & 54.85 \%\\
+  $so \wedge iy \wedge sr$:& 499 & 54.41 \%\\
+  $so \wedge sy \wedge sr$:& 544 & 59.32 \%\\
+  $     io \wedge iy$:& 486      & 52.99 \%\\
+  $     so \wedge iy$:& 506      & 55.17 \%\\
+  $     io \wedge sy$:& 509      & 55.50 \%\\
+  $     so \wedge sy$:& 551      & 60.08 \%\\
+  $     io \wedge ir$:& 525      & 57.25 \%\\
+  $     so \wedge ir$:& 578      & 63.03 \%\\
+  $     io \wedge sr$:& 623      & 67.93 \%\\
+  $     so \wedge sr$:& 719      & 78.40 \%\\
+  $          io$:& 629           & 68.59 \%\\
+  $          so$:& 730           & 79.60 \%\\
+\end{tabular}
+\caption{Absolute amount of annotation overlap.}\label{tab:1}
+
+\end{table}
+
+Table \ref{tab:1} shows the total amount of corresponding data in the individual
+markup. The table's labeling reads as follows: $s$ stands for a non-empty
+intersection between the two data sets (or prefix relation for the error types)
+and $i$ for equality.  $o,y,r$ denote the data type: $o$ stands for error
+tOkens, $y$ for the error's tYpe, and $r$ for the errors context (or
+\textit{tRigger}). Thus, $so\wedge sy\wedge ir$ is the number of all annotations
+that have a partial overlap on the error tokens $E$, a prefix match on the error
+type $\theta$, and identical error context.
+
+\section{Conclusion}
 
@@
 set of sequences of token indices.
+
+Annotating a corpus with this schema is quite laborious, and borders on
+impossibility without proper support from an annotation tool, such as the one we
+had to devise. During the process, however, our tool was constantly improved
+according to the annotators' ideas, and made the annotation process easier in
+the process. Partial automation and other features might increase annotation
+comfort even further.
+
+The taxonomy presented in \label{taxonomy} proved to be a little unwieldy. In
+particular, it did not clearly distinguish between annotating a target and
+annotating an error type. The next subsection proposes a refinement that might
+make the annotation process more precise.
+
+A part of speech tagger could aid the annotators and enhance the corpus data
+significantly. However, there are only a few reports on reliable part of speech
+tagging for learner language. \cite{luedeling} presents one such approach.
 

### 2009-04-09 text/paper.tex
Paper paper paper
SCISSSOSOOOSOSSORRSRSS!!!

@@
-\documentclass[12pt]{scrartcl}
+\documentclass[12pt]{article}
 \usepackage[utf8x]{inputenc}
@@
 \tableofcontents
-\abstract{I am an abstract!}
+\abstract{This paper presents an annotation method for errors in learner
+language that strives to be as compact and precise as possible, without
+sacrificing on generality or correctness. We develop contextualisation
+and underspecification methods for error annotations, based on a flexible annotation-driven markup.}
+
 \section{Introduction}\label{sec:intro}
 
-Literature overview. Bla Bla.
+Recently, there has been a growing interest in learner language corpora and
+learner errors analysis (citation). There is a need to
+identify errors in text, to make the automatic processing of those texts
+feasible\cite{nessel}. Also, the potential of learner errors for educational purposes and language
+acquisition insight has become increasingly recognized (citation), hence, many
+annotation schemes have been devised to identify and classify learner errors,
+such as \cite{negthesis}
 
-All of the annotation schemes mentioned above focus on the \textbf{text} itself,
-and will often even advise the annotator to \textit{modify} the underlying text
-with the annotation. We believe that a more error-centric annotation of the data
-can be beneficial.
+We propose a novel approach to annotating learner errors, based on the idea that
+an error is an inconsistency with its surroundings, rather .  This basic markup approach can be succesfully
+combined with any taxonomy for error ??classification. We have coupled it with a
+taxonomy for error classification and used the resulting scheme to annotate a
+subset of the NOce corpus (citation), achieving an inter-annotation agreement of
+???percent.
 
@@
 \centering
-\includegraphics[scale=.2]{taxonomy}
+\includegraphics[scale=.2]{taxonomy}\label{fig:tax}
 \caption{Error classification taxonomy.}
@@
 
-The source files to our corpus consisted of plain text files from the
+The source files for our corpus consisted of plain text files from the
 NOCE-corpus % FIXME cite
@@
 
+\appendix
+
+\section{Annotation Manual}
+  At any point, if you advance further down the taxonomy tree through a decision
+  you make later on, do not mark the error as the more general category, but
+  mark it as the more special category instead.
+
+\subsection{Steps in the annotation process}
+\begin{enumerate}
+\item Look at each token in turn, assuming all the rest is grammatical.
+\begin{itemize}
+  \item If the token is not spelled correctly, record a spelling error.
+  \item If an erroneous token $\tau_1$ is erroneous because it does not
+  harmonize with a token $\tau_2$
+  from its context,  mark $\tau_1$ as an error and $\tau_2$ as trigger.
+  \item If the utterance is ungrammatical, i.e. it does not constitute a part of
+  the language, mark it as a grammar mistake.
+  \item If the sentence is utterable under a certain condition, but this
+  condition is not met here, (i.e. forbidden by the general context of the
+  text) mark it as a contextual error.
+  \item Now, choose a subcategory from the branch you have chosen. Refer to the
+  graph in \ref{fig:tax}.
+  Try to be as specific as you can, otherwise, if not completely sure, choose
+  the upper node. Try to see in what way a token can be changed, as to
+  fit all the additional context. To find out which branch to choose, follow the
+  instructions under \ref{branches}
+  \item Repeat the process until all possible errors have been annotated, then
+  move to the next token.
+\end{itemize}
+
+  \item If the context is too erroneous and this erroneous context part affects
+  directly the immediate context of your token, try to find another solution,
+  because otherwise you will be building error hypotheses on errors. If it doesn't
+  (say, it is a spelling mistake), consider it as though it was correct while
+  looking for a directly interacting token as trigger.
+
+\item \textit{Local markup}: Keep it as local as possible, try to mark as
+triggers tokens that are grammatically dependent on each other in normal
+language production.
+
+\item Go from tokens up to other units, such as phrases, and be minimalistic, unless this makes you lose important information.
+
+\item Sometimes the error and/or the trigger will be a sequence. Take the
+minimal unit you can that doesn't lead to loss of information. Always choose
+the head of a constituent or compound word when it isn't the whole 
+constituent that plays a role in error formation.
+
+\ex.  The lovely, charming and amazing lady \textit{laugh} a lot
+
+Here, only take ``lady'' as a trigger. But in cases like \ref{granad}, your
+trigger should be ``to study'', and not only ``to'', because otherwise you'd lose
+information.  Maybe the learner doesn't have a problem with prepositions, but with infinitives.
+
+\ex. I go for to study in Granada.\label{granad}
+
+\item Avoid cascading and annotating errors based on corrected errors. Only look at the unaltered text.
+
+\item Annotate every possibility. Since we do not apply cascading rules, one can use weighting algorithms to prefer one error to another in the case of concurrent errors:
+
+\ex. \a.\textit{A} mobile \textbf{phones} can be very useful. 
+\b. \textbf{A} mobile \textit{phones} can be very useful.
+
+\item Only correct when necessary, do not be tempted to rephrase a construction,
+just because it would sound better or is more common, if there is nothing
+explicitly wrong with it.
+
+\item No style errors.
+
+\item No punctuation errors
+
+\item No capitalisation, only if it is something specific to English
+(Spelling$\to$Register), such as, say, capitalisation of proper nouns or months.
+
+\item If you have to choose between extremely similar possibilities, choose the
+one whose target hypothesis would have the least impact on the text in terms of
+phrasal constituents.
+
+\end{enumerate}
+
+\section{Branches in the taxonomy}\label{branches}
+
+\subsection{Collocation vs. Replace/Omission/Redundant}
+
+Verb and preposition collocations (i.e. where a certain verb forces a certain
+collocation) are to be marked as Collocations, not as an erroneous word. Words
+that occur with plural (\textit{one of}, \textit{between}) are also collocation
+bound, and an error in agreement is thus a collocation error.
+
+\subsection{Gerunds}
+
+Gerunds are nominalized noun forms.  Place mistakes related to the formation of
+gerunds in Morphology $\to$ Inflection, and errors related to the tense in which a
+gerund ought (not) to occur in Grammar $\to$ Verbtense.
+
+See also: Derivational Mistakes
+
+\subsection{Morphology}
+
+Morphological negation and plural formation mistakes often occur in English
+learner language. We tag them as Morphology and Morphology $\to$ Inflection instead
+of choosing a more specific category, but might extend our scheme in the future.
+
+\subsection{Derivational Mistakes}
+
+If a word is wrongly derived from another, (nominalization, suffixation,
+prefixation, etc\ldots) tag it as a morphological derivation mistake.
+
+\subsection{Animate/Inanimate}
+
+Is not covered by the scheme right now. Use Replace $\to$ Pronoun or similar. Might
+be added in the future.
+
+\subsection{Subject, Object, Predicate vs Pronoun, noun, verb.}
+
+If a whole constituent is missing or wrong, use the constituent classes. If it's
+one word out of the constituent, use the part of speech classes. Again: try to
+be as general as possible.
+
+\subsection{Number Agreement vs. Redundancy}
+
+\ex. a dogs
+
+\textit{a} is redundant, but because of number agreement. So you must annotate both,
+creating a circular dependency, where an erroneous \textit{dogs} has as error
+context \textit{a} and an erroneous \textit{a} has as context \textit{docs}
+
+\subsection{Compound nouns:}
+
+If a compound noun of two or more words is erroneous, annotate only the head noun, carrying the mistake.
+
+\subsection{Word order}
+\begin{itemize}
+\item \textit{Internal word order.} Words should be shuffled within the marked sequence. 
+\ex. Not only I must study $\to$ Not only must I study
+
+\item \textit{External word order.} Specified word needs to be moved somewhere
+else, its elements moving together as a unit.
+\ex. I the river like $\to$ I like the river
+
+\end{itemize}
+
+\subsection{Verb Tense}
+
+If the verb tense is wrongly formed, (\textit{he had having a nice time}), it's
+a grammatical error. Also, if there is a typical trigger word for verb tenses
+(such as \textit{since}, \textit{before}, \textit{if}) and the tense is wrong,
+it's a grammatical error. If the tense doesn't fit because of the broader
+context, it's a contextual mistake.
+
+\section{Source Code and XML Data}\label{sec:github}
+
+The complete code, as well as an up-to-date version of this paper can be found
+online at \texttt{www.github.com/adimit/ll-annotation}. The annotation schema's
+DTD is \texttt{AnnotationScheme.dtd} in the root directory of the repository.
+
 \end{document}

### 2009-04-09 text/paper.tex
bbbbbbbbbbbbbbbbbbbb

@@
 \usepackage{graphicx,linguex,natbib,aleks,amsmath,amsthm}
+\bibpunct{(}{)}{;}{a}{,}{,}
+\bibliographystyle{plainnat}
 \author{Magdalena Leshtanska \and Aleksandar Dimitrov}
@@
 Recently, there has been a growing interest in learner language corpora and
-learner errors analysis (citation). There is a need to
-identify errors in text, to make the automatic processing of those texts
-feasible\cite{nessel}. Also, the potential of learner errors for educational purposes and language
-acquisition insight has become increasingly recognized (citation), hence, many
-annotation schemes have been devised to identify and classify learner errors,
-such as \cite{negthesis}
+learner errors analysis \cite{nessel}. There is a need to identify errors in
+text produced by second language learners, to make the automatic processing of
+those texts feasible (ibid.). Also, the potential of learner errors for
+educational purposes and language acquisition insight has become increasingly
+recognized (ibid.), hence, many annotation schemes have been devised to
+identify and classify learner errors, such as \cite{negrillo}, \cite{CLC},
+\cite{ludeling}
 
 We propose a novel approach to annotating learner errors, based on the idea that
-an error is an inconsistency with its surroundings, rather .  This basic markup approach can be succesfully
-combined with any taxonomy for error ??classification. We have coupled it with a
-taxonomy for error classification and used the resulting scheme to annotate a
-subset of the NOce corpus (citation), achieving an inter-annotation agreement of
-???percent.
+an error is an inconsistency with its surroundings. This basic markup approach
+can be  combined with a taxonomy for error classification. We
+have coupled it with learner error  classification system and used the resulting
+scheme to annotate a subset of the NOCE corpus (\cite{negrilloph}), a corpus of
+beginning to intermediate English learners of Spanish.
 
-Firstly, every annotation scheme that operates directly on the text or spans
-over a piece of text will run into two kinds of problems:
-
-\begin{itemize}
-\item \textit{Interleaving annotations} occur when an error doesn't end before
-another one begins. Given the tokens $\tau_1 \tau_2 \tau_3$, and two errors
-$\eta_1$ and
-$\eta_2$ ranging over tokens $\tau_1$, $\tau_2$ and $\tau_2$ $\tau_3$ respectively, the
-resulting markup will be confusing or outright impossible to read: $(\eta_1)\tau_1
-(\eta_2)\tau_2(/\eta_1)\tau_3(/\eta_2)$. This is particularly a problem with XML,
-since the specification\footnotemark explicitly disallows interleaving
-markup.
-\item \textit{Greedy annotation} covers tokens entirely uninvolved in the
-\textit{``production''} of an error. If of the token sequence $\tau_1 \tau_2
-\tau_3$ only $\tau_1$
-and $\tau_3$ are erroneous,
-marking the entire token sequence with an error annotation would falsely accuse
-the
-otherwise completely innocent token $\tau_2$.
-\end{itemize}
-\footnotetext{Located at \texttt{http://www.w3.org/TR/REC-xml/}}
-
-Based on these assumptions, we decided to decouple the \textit{error markup}
-from the \textit{corpus data}. Specifically, to our annotation method, the
-errors and the text are two entirely different data structures. Every
-particular error can reference tokens within the corpus using a \textbf{key},
-similar to the way modern Relational Data Bases reference their data. This makes
-the index more accessible and easier to maintain and eliminates both problems
-above, because every single annotation can be completely independent from all
-other annotations.
+Section \ref{sec:scheme} will introduce the general idea behind our annotation
+scheme and elaborate on more technical details. Section 3  gives a brief
+overview of real world performance of our method, as far as this is indeed
+possible. An annotation manual is contained in Appendix A.
 
@@
 
-In order to identify errors occurring in learner language, one has to first
-classify them. However, even then, identifying which class a particular error
-belongs to is in no way an easy job. Devising an exhaustive taxonomy of errors
-that can appear in natural language seems a daunting task, since science has so
-far failed to regularize what \textit{is} a valid utterance of a language.
-Therefore, our taxonomy will try to do some things only, and do them well. In
-particular, we will not cover punctuation mistakes, since their possible
-corrections can have cascading effects on the rest of the sentences and suddenly
-give rise to all kinds of new errors.
+Devising an exhaustive taxonomy of errors that can appear in natural language
+seems a daunting task, since science has so far failed to regularize what
+\textit{is} a valid utterance of a language.  Therefore, our annotation scheme
+strives  to  do some things only, and do them well.
 
@@
 
-For our test taxonomy, we used the taxonomy described in \cite{negrillo} as a
+For our test taxonomy, we used the taxonomy described in \cite{negrilloph} as a
 base, and enhanced it by several concepts from the CLC annotation scheme, as
@@
   \item \textit{Style Errors.} Style is too soft a concept. Errors in style are
-  notoriously hard to peg and reliably quantify. Moreover, the NOCE corpus
+  notoriously hard to peg and reliably quantify. Moreover, our corpus
   contained data mostly from beginning English learners, where style errors are
@@
 
-The source files for our corpus consisted of plain text files from the
-NOCE-corpus % FIXME cite
-of beginning to intermediate Spanish learners of English. Our corpus
-format is defined in \ref{def:errorformat}
+The context mechanism in our annotation scheme can increase the complexity of the annotation
+process drastically. Since our annotation technique requires us to build a
+non-hierarchical graph structure over our input\footnotemark, a traditional
+sequential markup scheme scheme that operates directly on the text or spans
+over a piece of text will run into at least two kinds of problems:
+
+\footnotetext{Relating words and their dependencies more like a net than as a
+hierarchically nested span over a character sequence}
+
+\begin{itemize}
+\item \textit{Interleaving annotations} occur when an error doesn't end before
+another one begins. Given the tokens $\tau_1 \tau_2 \tau_3$, and two errors
+$\eta_1$ and
+$\eta_2$ ranging over tokens $\tau_1$, $\tau_2$ and $\tau_2$ $\tau_3$ respectively, the
+resulting markup will be confusing or outright impossible to read: $(\eta_1)\tau_1
+(\eta_2)\tau_2(/\eta_1)\tau_3(/\eta_2)$. This is particularly a problem with XML,
+since the specification\footnotemark explicitly disallows interleaving
+markup.
+\item \textit{Greedy annotation} covers tokens entirely uninvolved in the
+\textit{``production''} of an error. If of the token sequence $\tau_1 \tau_2
+\tau_3$ only $\tau_1$
+and $\tau_3$ are erroneous,
+marking the entire token sequence with an error annotation would falsely accuse
+the
+otherwise completely innocent token $\tau_2$.
+\end{itemize}
+\footnotetext{Located at \texttt{http://www.w3.org/TR/REC-xml/}}
+
+Based on these assumptions, we decided to decouple the \textit{error markup}
+from the \textit{corpus data}. Specifically, to our annotation method, the
+errors and the text are two entirely different data structures. Every particular
+error can reference tokens within the corpus using a \textbf{key}, similar to
+the way modern Relational Data Bases reference their data. This makes the index
+more accessible and easier to maintain and eliminates both problems above,
+because every single annotation can be completely independent from all other
+annotations.  The source files for our corpus consisted of plain text files from
+the NOCE-corpus of beginning to intermediate Spanish learners of English. Our
+corpus format is defined in \ref{def:errorformat}.
 
@@
 \subsection{Limitations}
-Some elements, for example word order errors have enforced context, but it is hard to express via our schema, so we don't do it. 
-No correct positions specified by word order errors - also hard to express.
-One could assign them a target position after an existing token, but then one must also keep track of other "reorderings" by multiple word order errors - I would like beer some to drink - here, both erroneous units - some and to drink should be placed between like and beer, but one must take care to order them correctly.
-Also, as our schema tags all possible errors, in the case of interchangeable errors (cf XXX), one cannot be certain if the error annotated is really the one, made by the learner, and not the other possible one.
-Most  importantly, there are sometimes more than one places where a word sequence can fit in.
+Word order errors still pose a particular complication even to a non-sequential
+scheme such as ours. We are not able to account for the complex dependencies
+occurring in some word order mistakes simply by a dependency-graph.
 
-Errors within a token cannot be analyzed – for example, agglutination, because a token is our smallest unit.
+Since our minimal atomic unit is a token, we depend on a tokenization process.
+Furthermore, we cannot account for mistakes that occur on a scale smaller than a
+token.
 
@@
 on a particular data set by two or more annotators are compared using
-quantitative methods.  \cite{ap2008} give an overview of currently employed
+quantitative methods.  \cite{artstein} give an overview of currently employed
 methods.
@@
 
-\cite{ap2008} briefly discuss unitization and go on to note that it has thus far
+\cite{artstein} briefly discuss unitization and go on to note that it has thus far
 not been exhaustively researched. Even more importantly, they explicitly comment
@@
 measure in the corpus linguistic literature, $\alpha_U$, presented in
-\cite{krip1995}.  Apart from being untested, $\alpha_U$ also seems to have
+\cite{krip}.  Apart from being untested, $\alpha_U$ also seems to have
 problems with overlapping markup both in the text, and between annotation sets
@@
 After assessing the quality of our data, we reached
-the conclusion that the error format described in \ref{sec:errorformat}
+the conclusion that the error format described in \ref{sec:taxonomy}
 might benefit from several refinements. Adding a field for part of speech tags
@@
 significantly. However, there are only a few reports on reliable part of speech
-tagging for learner language. \cite{luedeling} presents one such approach.
+tagging for learner language. \cite{ludeling} presents one such approach.
 
@@
 
-\subsection{Compound nouns:}
+\subsection{Compound nouns}
 
 If a compound noun of two or more words is erroneous, annotate only the head noun, carrying the mistake.
+
+\subsection{Sequences}
+For both triggers and erroneous tokens, it is best to use a single token. In
+case of a phrase, only take the keyword and at most any relevant information
+bearing token. In such a case, do not care about order.
+
+Still, fixed expressions, such as  Collocations or Idiom can be counted as one
+unit. In such a case, annotate the whole phrase, even if it contains erroneous
+tokens. Only do not annotate the target if it is within the idiom.
 
@@
 
-\section{Source Code and XML Data}\label{sec:github}
+\section{Source Code and XML Data}\label{appendix:github}
 
@@
 
+\bibliography{bib}
+
 \end{document}

### 2009-04-09 text/paper.tex
aaaaaaaaaaaaaaaaaaaaaaaaaaaaaa

@@
 Section \ref{sec:scheme} will introduce the general idea behind our annotation
-scheme and elaborate on more technical details. Section 3  gives a brief
-overview of real world performance of our method, as far as this is indeed
-possible. An annotation manual is contained in Appendix A.
+scheme and elaborate on more technical details. Section 4 gives a brief overview
+of real world performance of our method. An annotation manual is contained in
+Appendix A.
 
@@
 these problems.
+
+Firstly, every annotation scheme that operates directly on the text or spans
+over a piece of text will run into two kinds of problems:
+
+\begin{itemize}
+\item \textit{Interleaving annotations} occur when an error doesn't end before
+another one begins. Given the tokens $\tau_1 \tau_2 \tau_3$, and two errors
+$\eta_1$ and
+$\eta_2$ ranging over tokens $\tau_1$, $\tau_2$ and $\tau_2$ $\tau_3$ respectively, the
+resulting markup will be confusing or outright impossible to read: $(\eta_1)\tau_1
+(\eta_2)\tau_2(/\eta_1)\tau_3(/\eta_2)$. This is particularly a problem with XML,
+since the specification\footnotemark explicitly disallows interleaving
+markup.
+\item \textit{Greedy annotation} covers tokens entirely uninvolved in the
+\textit{``production''} of an error. If of the token sequence $\tau_1 \tau_2
+\tau_3$ only $\tau_1$
+and $\tau_3$ are erroneous,
+marking the entire token sequence with an error annotation would falsely accuse
+the
+otherwise completely innocent token $\tau_2$.
+\end{itemize}
+\footnotetext{Located at \texttt{http://www.w3.org/TR/REC-xml/}}
+
+Based on these assumptions, we decided to decouple the \textit{error markup}
+from the \textit{corpus data}. Specifically, to our annotation method, the
+errors and the text are two entirely different data structures. Every
+particular error can reference tokens within the corpus using a \textbf{key},
+similar to the way modern Relational Data Bases reference their data. This makes
+the index more accessible and easier to maintain and eliminates both problems
+above, because every single annotation can be completely independent from all
+other annotations.
 
@@
 
-The context mechanism in our annotation scheme can increase the complexity of the annotation
-process drastically. Since our annotation technique requires us to build a
-non-hierarchical graph structure over our input\footnotemark, a traditional
-sequential markup scheme scheme that operates directly on the text or spans
-over a piece of text will run into at least two kinds of problems:
-
-\footnotetext{Relating words and their dependencies more like a net than as a
-hierarchically nested span over a character sequence}
-
-\begin{itemize}
-\item \textit{Interleaving annotations} occur when an error doesn't end before
-another one begins. Given the tokens $\tau_1 \tau_2 \tau_3$, and two errors
-$\eta_1$ and
-$\eta_2$ ranging over tokens $\tau_1$, $\tau_2$ and $\tau_2$ $\tau_3$ respectively, the
-resulting markup will be confusing or outright impossible to read: $(\eta_1)\tau_1
-(\eta_2)\tau_2(/\eta_1)\tau_3(/\eta_2)$. This is particularly a problem with XML,
-since the specification\footnotemark explicitly disallows interleaving
-markup.
-\item \textit{Greedy annotation} covers tokens entirely uninvolved in the
-\textit{``production''} of an error. If of the token sequence $\tau_1 \tau_2
-\tau_3$ only $\tau_1$
-and $\tau_3$ are erroneous,
-marking the entire token sequence with an error annotation would falsely accuse
-the
-otherwise completely innocent token $\tau_2$.
-\end{itemize}
-\footnotetext{Located at \texttt{http://www.w3.org/TR/REC-xml/}}
-
-Based on these assumptions, we decided to decouple the \textit{error markup}
-from the \textit{corpus data}. Specifically, to our annotation method, the
-errors and the text are two entirely different data structures. Every particular
-error can reference tokens within the corpus using a \textbf{key}, similar to
-the way modern Relational Data Bases reference their data. This makes the index
-more accessible and easier to maintain and eliminates both problems above,
-because every single annotation can be completely independent from all other
-annotations.  The source files for our corpus consisted of plain text files from
-the NOCE-corpus of beginning to intermediate Spanish learners of English. Our
-corpus format is defined in \ref{def:errorformat}.
+The source files for our corpus consisted of plain text files from the
+NOCE-corpus (\cite{negrilloph})
+of beginning to intermediate Spanish learners of English. Our corpus
+format is defined in \ref{def:errorformat}
 
@@
 \subsection{Limitations}
-Word order errors still pose a particular complication even to a non-sequential
-scheme such as ours. We are not able to account for the complex dependencies
-occurring in some word order mistakes simply by a dependency-graph.
+Some elements, for example word order errors have enforced context, but it is hard to express via our schema, so we don't do it. 
+No correct positions specified by word order errors - also hard to express.
+One could assign them a target position after an existing token, but then one must also keep track of other "reorderings" by multiple word order errors - I would like beer some to drink - here, both erroneous units - some and to drink should be placed between like and beer, but one must take care to order them correctly.
+Also, as our schema tags all possible errors, in the case of interchangeable errors (cf XXX), one cannot be certain if the error annotated is really the one, made by the learner, and not the other possible one.
+Most  importantly, there are sometimes more than one places where a word sequence can fit in.
 
-Since our minimal atomic unit is a token, we depend on a tokenization process.
-Furthermore, we cannot account for mistakes that occur on a scale smaller than a
-token.
+Errors within a token cannot be analyzed – for example, agglutination, because a token is our smallest unit.
 

### 2009-04-09 text/paper.tex
Finally?

--- 2009-04-09 text/paper.tex
@@
-\documentclass[12pt]{article}
+\documentclass{article}
 \usepackage[utf8x]{inputenc}
@@
 We propose a novel approach to annotating learner errors, based on the idea that
-an error is an inconsistency with its surroundings. This basic markup approach
+an error is an inconsistency with its context. This basic markup approach
 can be  combined with a taxonomy for error classification. We
@@
 scheme to annotate a subset of the NOCE corpus (\cite{negrilloph}), a corpus of
-beginning to intermediate English learners of Spanish.
+beginner to intermediate English learners of Spanish.
 
@@
 overview of real world performance of our method, as far as this is indeed
-possible. An annotation manual is contained in Appendix A.
+possible. Appendix A contains the annotation manual.
 
@@
 
-Our annotation scheme presents three distinct kinds of errors:
+Our annotation scheme presents three distinct kinds of errors based on a
+definition of an error as an inconsistency with the context:
 
@@
   We chose to ignore such errors, since they are not a sign of weakness in the
-  target language but likely pure by chance mistakes.
+  target language but likely pure by chance mistakes or general incompetence.
   \item \textit{Style Errors.} Style is too soft a concept. Errors in style are
   notoriously hard to peg and reliably quantify. Moreover, our corpus
-  contained data mostly from beginning English learners, where style errors are
+  contained data mostly from beginner English learners, where style errors are
   not as important, evident or relevant.
@@
 because every single annotation can be completely independent from all other
-annotations.  The source files for our corpus consisted of plain text files from
-the NOCE-corpus of beginning to intermediate Spanish learners of English. Our
-corpus format is defined in \ref{def:errorformat}.
+annotations. Our corpus format is defined in \ref{def:errorformat}.
 
@@
 common task in the annotation of learner language is \textit{delineating}
-the extent of a certain mistake.
+the extent of a certain mistake. This has been called \textbf{unitization} by
+\cite{krip}.
 
@@
 on the unknown status of the validity of the only inter-annotator agreement
-measure in the corpus linguistic literature, $\alpha_U$, presented in
+measure in the literature, $\alpha_U$, presented in
 \cite{krip}.  Apart from being untested, $\alpha_U$ also seems to have
@@
 markup. The table's labeling reads as follows: $s$ stands for a non-empty
-intersection between the two data sets (or prefix relation for the error types)
-and $i$ for equality.  $o,y,r$ denote the data type: $o$ stands for error
-tOkens, $y$ for the error's tYpe, and $r$ for the errors context (or
-\textit{tRigger}). Thus, $so\wedge sy\wedge ir$ is the number of all annotations
-that have a partial overlap on the error tokens $E$, a prefix match on the error
-type $\theta$, and identical error context.
+intersection between the two data sets' fields. For the error types, $s$ was
+defined to be a prefix operator instead, accounting for common paths along the
+taxonomy tree.  $i$ stands for equality.  $o,y,r$ denote the data type: $o$
+stands for error tOkens, $y$ for the error's tYpe, and $r$ for the errors
+context (or \textit{tRigger}). Thus, $so\wedge sy\wedge ir$ is the number of all
+annotations that have a partial overlap on the error tokens $E$, a prefix match
+on the error type $\theta$, and identical error context.
+
+Having a look at the data presented in table \ref{tab:1}, the figures do not
+seem to be in strong support of our efforts thus far, although there is little
+grounds for comparison with other similar data. Nevertheless the figures may
+indeed allow partial assessment of the data.  Less than half of the annotations
+were completely equivalent. Interestingly, while the discrepancy between a
+strict and a fuzzy match on the token sequences alone is above 10\%, it is very
+low when we demand matches on the rest of the data structure, too. This probably means
+that we have an overhead of unclear cases, which are currently hard to catch,
+whereas on the other hand we have a clear cut kernel of reliable rules, too.
+
+Upon inspecting the corpus data in greater detail, we found out that it was
+sometimes hard to decide whether a particular error was a contextual or
+grammatic error. Moreover, often the categories within the grammatic error
+branch were not clear enough. It seems that our merger of the CLC and Free Text
+introduced some confusion with certain categories. Particularly, the
+\textit{omission}, \textit{replace}, and \textit{redundancy} cases would at
+times `steal' hits from the other categories due to their very general
+nomenclature. We have revised the annotation manual to specifically mention that
+all other grammatical error categories take precedence over these three.
+Unfortunately, this seems to be a suboptimal solution. The next section contains
+a proposal for a much cleaner solution.
+
+In revising the annotation scheme, the comment field of our annotations proved
+to be quite useful. Both annotators commented frequently that they lacked
+categories for adverbs and gerunds. The former has been integrated in to the
+scheme already, and we might add a gerund field in the future, too.
+
+Another very unclear point in the original manual was the minimalization of
+error spans constraint, which advises the annotator to only select the smallest
+possible amount of tokens, even when that means splitting up compound nouns.
+Also, both annotators reacted differently to situations where a sequence had to
+be marked up. This spawned a new subsection in our annotation manual as well.
 

### 2009-04-09 text/paper.tex
Final!

@@
 sometimes hard to decide whether a particular error was a contextual or
-grammatic error. Moreover, often the categories within the grammatic error
+grammatic error. Moreover, the categories within the grammatic error
 branch were not clear enough. It seems that our merger of the CLC and Free Text
@@
 \textit{omission}, \textit{replace}, and \textit{redundancy} cases would at
-times `steal' hits from the other categories due to their very general
-nomenclature. We have revised the annotation manual to specifically mention that
+times `steal' hits from the other categories due to their general
+nomenclature. We revised the annotation manual to specifically mention that
 all other grammatical error categories take precedence over these three.
@@
 
-Another very unclear point in the original manual was the minimalization of
-error spans constraint, which advises the annotator to only select the smallest
+However, the error type taxonomy was not our primary concern. In theory, any
+taxonomy would fit. We believe that taxonomies with a finer granularity might be
+preferable. Our main interest was the agreement on error boundary detection,
+error context detection and the three basic error types.  A very unclear point
+in that context with respect to the original manual was the minimalization of
+token spans constraint, which advises the annotator to only select the smallest
 possible amount of tokens, even when that means splitting up compound nouns.
 Also, both annotators reacted differently to situations where a sequence had to
-be marked up. This spawned a new subsection in our annotation manual as well.
+be marked up. We discovered that this had quite an effect on the consistency of
+our annotations and decided to add a new subsection in our annotation manual
+about sequences.
 
@@
 
-Moreover, we came to the conclusion that defining the error context as a set of
+Moreover, we saw that defining the error context as a set of
 tokens might be misleading or at least difficult to understand in case the error
@@
 had to devise. During the process, however, our tool was constantly improved
-according to the annotators' ideas, and made the annotation process easier in
+according to the annotators' ideas, and made the annotation task easier in
 the process. Partial automation and other features might increase annotation
